% !TeX program = pdflatex
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{setspace}
\onehalfspacing

\usepackage{amsmath,amssymb,amsfonts,mathtools,bm,booktabs,tabularx,array}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[round,authoryear]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Rp}{\mathbb{R}_{+}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\Id}[1]{\mat I_{#1}}
\newcommand{\zeros}[1]{\vect 0_{#1}}
\newcommand{\ones}[1]{\vect 1_{#1}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\Normal}{\mathcal{N}}
\newcommand{\TN}{\mathcal{N}^{+}}
\newcommand{\IG}{\operatorname{IG}}
\newcommand{\GIG}{\operatorname{GIG}}
\newcommand{\Exp}{\operatorname{Exp}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\AL}{\operatorname{AL}}
\newcommand{\exAL}{\operatorname{exAL}}
\newcommand{\RHS}{\operatorname{RHS}}
\newcommand{\elbo}{\mathcal{L}}
\newcommand{\logit}{\operatorname{logit}}
\newcommand{\expit}{\operatorname{expit}}
\newcommand{\pkg}[1]{\textsf{#1}}
\newcommand{\TableStyle}{\footnotesize\setlength{\tabcolsep}{4pt}\renewcommand{\arraystretch}{1.08}}

\newcounter{suppalgorithm}
\renewcommand{\thesuppalgorithm}{S\arabic{suppalgorithm}}
\newenvironment{suppalgorithm}[1]{%
  \refstepcounter{suppalgorithm}%
  \par\medskip\noindent\textbf{Algorithm~\thesuppalgorithm. #1.}\par\nobreak
}{\par\medskip}

\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thesubsection}{S\arabic{section}.\arabic{subsection}}
\renewcommand{\thesubsubsection}{S\arabic{section}.\arabic{subsection}.\arabic{subsubsection}}

\title{Supplementary Material for ``Bayesian Quantile Deep Echo State Networks''}
\author{Antonio De Leon \and Raquel Prado \and Bruno Sans\'o}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\noindent
Conditional on fixed reservoir features, Q--DESN is a high-dimensional
Bayesian quantile readout model under the asymmetric Laplace (AL) or extended
asymmetric Laplace (exAL) working likelihood. The augmented posterior target
includes the readout coefficients, likelihood scale and asymmetry, auxiliary
mixture variables, and shrinkage parameters. The derivations cover complete-data
posterior targets for the default ridge prior and the Nishimura--Suchard
regularized horseshoe (\texttt{rhs\_ns}) shrinkage alternative, posterior
kernels, Markov chain Monte Carlo (MCMC) full conditionals, mean-field
variational Bayes coordinate updates, the Laplace--Delta treatment of
non-conjugate blocks, evidence lower bound (ELBO) terms, and the joint
quantile-vector QVP--RHS extension.
It also records the relaxed quantile regression DESN (RQR--DESN)
interval-readout extension as a generalized-Bayes target over two fixed-feature
root readouts.
The notation is aligned with the main article.
\end{abstract}

\section{Scope and Relation to the Main Article}
\label{sec:supp_scope}

The main article introduces Q--DESN, describes the fixed deep reservoir feature
map in the notation and model-specification sections, and records the simulation
design in the simulation section. Once
the reservoir weights, input weights, reducers, leak rates, and washout
convention are fixed, the likelihood depends on the data only through a
deterministic design matrix
\(\mat X=[\vect x_1^\top;\ldots;\vect x_T^\top]\). Thus the
Q--DESN readout is a static Bayesian quantile regression problem with a
high-dimensional nonlinear design. Consequently, the static AL and exAL
regression derivations used in the
\pkg{exdqlm} supplement \citep{barata2021_flex_quantile} can be adapted to the
Q--DESN readout by replacing ordinary covariate rows with DESN feature rows.

The posterior target, MCMC sampler, CAVI updates, and ELBO are stated so that
the computations are recoverable without consulting implementation code. Ridge
is the default coefficient prior. The main article uses \(\RHS\) for the
regularized horseshoe readout prior, while \texttt{rhs\_ns} denotes the
Nishimura--Suchard product representation used for closed-form scale updates.
The main article states the posterior target and algorithms; this supplement
contains the recoverable full conditionals, variational factors, and ELBO
terms that support those algorithms.
Table~\ref{tab:supp_roadmap} gives a navigation map.

\begin{table}[!htb]
\centering
\begingroup
\TableStyle
\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.25\textwidth}
>{\raggedright\arraybackslash}p{0.24\textwidth}
>{\raggedright\arraybackslash}X@{}}
\toprule
Topic & Supplement location & Role in Q--DESN \\
\midrule
Reservoir and readout notation &
Sections~\ref{sec:desn_features}--\ref{sec:model_joint} &
Defines the deterministic design matrix and the likelihood after feature
construction. \\
AL and exAL complete-data targets &
Section~\ref{sec:model_joint} &
States the joint distribution for likelihood, latent variables, and priors. \\
Default ridge and \texttt{rhs\_ns} priors &
Sections~\ref{sec:priors} and \ref{sec:rhs_ns} &
Gives the default Gaussian readout prior and the adaptive shrinkage alternative
used in the simulation tables. \\
MCMC full conditionals &
Section~\ref{sec:mcmc} &
Derives the readout, latent-variable, scale, asymmetry, and shrinkage updates. \\
Variational Bayes and Laplace--Delta updates &
Section~\ref{sec:cavi} &
Derives the mean-field factors, closed-form updates, and Laplace--Delta block
used by the variational Bayes algorithm. \\
ELBO &
Section~\ref{sec:elbo} &
Gives the exact and approximate terms monitored by the VB algorithm. \\
Joint quantile-vector extension &
Section~\ref{sec:joint_qdesn_supp} &
Derives the stacked exAL likelihood, QVP--RHS prior, MCMC blocks, VB--LD
factorization, ELBO terms, and crossing diagnostics for the joint
multi-quantile readout. \\
RQR--DESN interval readout &
Section~\ref{sec:rqr_desn_supp} &
Defines the two-root generalized-Bayes interval target, pseudo-AL
augmentation, and MCMC block structure for the supplementary interval-readout
extension. \\
Posterior prediction and synthesis &
Section~\ref{sec:forecast_synthesis} &
Links the fitted quantile readout to working-likelihood forecast simulation and
multi-quantile synthesis. \\
GloFAS discrepancy-calibration model &
Section~\ref{sec:glofas_discrepancy} &
Expresses the single-forecast-system application as an augmented
Q--DESN readout with reference and GloFAS-specific likelihood blocks. \\
\bottomrule
\end{tabularx}
\endgroup
\caption{Supplement roadmap for the Q--DESN posterior target and algorithms.}
\label{tab:supp_roadmap}
\end{table}

\section{Distributional Conventions}
\label{sec:dist_conventions}

The posterior kernels below use the following distributional
parameterizations. These conventions are inherited by the MCMC, CAVI, and ELBO
derivations below; in particular, all inverse-gamma distributions use a rate in
the reciprocal scale.

The inverse-gamma (IG) parameterization is
\[
p(x\mid a,b)=\frac{b^a}{\Gamma(a)}x^{-a-1}\exp(-b/x),\qquad x>0,
\]
and the generalized inverse Gaussian (GIG) parameterization
\[
p(x\mid \lambda,\chi,\psi)
\propto x^{\lambda-1}\exp\left\{-\frac12(\chi/x+\psi x)\right\},
\qquad x>0.
\]
The notation \(\TN(m,V)\) denotes a Normal distribution with mean \(m\) and
variance \(V\), truncated to \((0,\infty)\). The functions \(\Phi\) and
\(\phi\) denote the standard Normal distribution and density functions.

For a fixed target quantile \(p_0\in(0,1)\), the AL constants are
\[
A_0=\frac{1-2p_0}{p_0(1-p_0)},\qquad
B_0=\frac{2}{p_0(1-p_0)}.
\]
The notation \(\exAL\) denotes the quantile-fixed generalized asymmetric
Laplace (GAL) distribution of \citet{yan2025new}. The exAL likelihood
introduces an asymmetry parameter \(\gamma\in(L,U)\), where \(L\) and
\(U\) are determined by the quantile-identifying construction. Define
\[
g(\gamma)=2\Phi(-|\gamma|)\exp(\gamma^2/2),\qquad
p_\gamma=\ind\{\gamma<0\}
+\frac{p_0-\ind\{\gamma<0\}}{g(\gamma)}.
\]
Then
\[
A_\gamma=\frac{1-2p_\gamma}{p_\gamma(1-p_\gamma)},\quad
B_\gamma=\frac{2}{p_\gamma(1-p_\gamma)},\quad
C_\gamma=\{\ind\{\gamma>0\}-p_\gamma\}^{-1},\quad
D_\gamma=C_\gamma|\gamma|.
\]
The main article denotes the same exAL shift coefficient by
\(\lambda(\gamma)=C(\gamma)|\gamma|\); the supplement uses \(D_\gamma\)
to reduce visual clutter in posterior kernels.
When \(\gamma=0\) and the \(s_t\) latents are omitted, the exAL hierarchy
reduces to the AL working likelihood with constants \(A_0,B_0\).

\section{DESN Feature Map and Static Readout}
\label{sec:desn_features}

The Q--DESN readout inherits the feature map defined in the main article. The
reservoir weights, input weights, reducers, leak rates, and washout convention
are fixed before posterior inference, so the only object carried into the static
readout model is the design matrix \(\mat X\). Scalar responses are denoted by
lowercase \(y_t\), both in likelihood statements and posterior kernels.

Let \(\vect u_t=(y_{t-1},\ldots,y_{t-m},\vect z_t^\top)^\top\) collect response
lags and exogenous covariates. For layers \(d=1,\ldots,D\), Q--DESN draws sparse
reservoir matrices \(\mat W_d\), input matrices \(\mat W^{\mathrm{in}}_d\), and
optional reducers \(\mat Q_d\), \(d=1,\ldots,D-1\), once and keeps them fixed.
With layer-specific leak rates \(\alpha_d\), activation \(f\), and spectral
radius targets \(\rho_d\),
the layer states satisfy
\begin{align}
\vect h_{t,1}
&=(1-\alpha_1)\vect h_{t-1,1}
+\alpha_1 f\{\bar{\mat W}_1\vect h_{t-1,1}+\mat W^{\mathrm{in}}_1\vect u_t\},
\label{eq:supp_desn_layer1}\\
\vect h_{t,d}
&=(1-\alpha_d)\vect h_{t-1,d}
+\alpha_d f\{\bar{\mat W}_d\vect h_{t-1,d}
+\mat W^{\mathrm{in}}_d\tilde{\vect h}_{t,d-1}\},\qquad d\ge2,
\label{eq:supp_desn_layerd}\\
\tilde{\vect h}_{t,d-1}&=\mat Q_{d-1}\vect h_{t,d-1},
\end{align}
where \(\bar{\mat W}_d=\rho_d\mat W_d/\rho(\mat W_d)\). The readout feature
vector is
\begin{equation}
\tilde{\vect x}_t
=\{ \vect h_{t,D}^\top,\ k(\tilde{\vect h}_{t,1})^\top,\ldots,
k(\tilde{\vect h}_{t,D-1})^\top,\ \vect u_t^\top\}^\top,
\qquad
\vect x_t=(1,\tilde{\vect x}_t^\top)^\top.
\label{eq:supp_feature_stack}
\end{equation}
After a fixed washout period, the reservoir quantities in
\eqref{eq:supp_feature_stack} are treated as deterministic. Let
\(\vect\beta=(\beta_0,\beta_1,\ldots,\beta_r)^\top\) and
\[
\mat X=[\vect x_1^\top;\ldots;\vect x_T^\top]\in\R^{T\times(r+1)},\qquad
\mu_t=\vect x_t^\top\vect\beta .
\]
All posterior inference is conditional on \(\mat X\).

\section{Q--DESN Likelihoods, Priors, and Posterior Targets}
\label{sec:model_joint}

The complete-data posterior targets below are used by both computational
schemes. The AL and exAL likelihoods are working likelihoods for a
fixed target quantile, while the auxiliary variables provide equivalent
mixture representations for computation.

\subsection{AL and exAL Augmented Likelihoods}

The exAL Q--DESN observation hierarchy is
\begin{align}
v_t\mid\sigma &\sim \Exp(\text{rate}=1/\sigma),&
s_t&\sim\TN(0,1),\\
y_t\mid\vect\beta,\sigma,\gamma,v_t,s_t,\mat X
&\sim
\Normal\left(
\vect x_t^\top\vect\beta+\sigma D_\gamma s_t+A_\gamma v_t,
\sigma B_\gamma v_t
\right).
\label{eq:supp_exal_hierarchy}
\end{align}
The marginal \(p_0\)-quantile of \(y_t\mid\vect x_t\) is
\(\vect x_t^\top\vect\beta\). The AL special case is
\begin{align}
v_t\mid\sigma &\sim \Exp(\text{rate}=1/\sigma),\\
y_t\mid\vect\beta,\sigma,v_t,\mat X
&\sim
\Normal\left(
\vect x_t^\top\vect\beta+A_0v_t,
\sigma B_0v_t
\right).
\label{eq:supp_al_hierarchy}
\end{align}

\subsection{Coefficient Priors}
\label{sec:priors}

The coefficient prior regularizes the high-dimensional readout after the
reservoir features have been fixed. The ridge prior is the default Gaussian
baseline. The \texttt{rhs\_ns} prior is the adaptive global-local shrinkage
alternative; its product representation keeps the local, global, and slab scale
updates in closed form.

For the ridge readout prior, let
\begin{equation}
\vect\beta\sim\Normal(\vect b_\beta,\mat V_\beta),
\qquad
\mat P_\beta=\mat V_\beta^{-1}.
\label{eq:supp_ridge_prior}
\end{equation}
Applications usually keep the intercept weakly regularized and use a
diagonal slope penalty, while \eqref{eq:supp_ridge_prior} allows a general
positive definite prior covariance.

For the \texttt{rhs\_ns} prior, the intercept is independent,
\(\beta_0\sim\Normal(b_0,V_0)\), and the non-intercept coefficients are
represented through the
Nishimura--Suchard-style regularized horseshoe construction
\citep{CarvalhoPolsonScott2010HS,PiironenVehtari2017RHS,NishimuraSuchard2023SSS}.
For \(j=1,\ldots,r\),
\begin{equation}
p_{\mathrm{ns}}(\beta_j,\lambda_j\mid\tau,\zeta)
\propto
\exp\left(-\frac{\beta_j^2}{2\tau^2\lambda_j^2}\right)
\exp\left(-\frac{\beta_j^2}{2\zeta^2}\right)
\lambda_j^{-1}(1+\lambda_j^2)^{-1}.
\label{eq:supp_ns_joint}
\end{equation}
Equation~\eqref{eq:supp_ns_joint} is the marginal joint description after the
half-Cauchy auxiliary variable for the local scale has been integrated out. The
MCMC and VB derivations use the equivalent augmented product
\[
\Normal(\beta_j\mid0,\tau^2\lambda_j^2)\,
\Normal(0\mid\beta_j,\zeta^2)
\]
together with inverse-gamma auxiliaries for \(\lambda_j^2\), \(\nu_j\),
\(\tau^2\), and \(\xi\). This augmented product representation is the one used
in the complete-data joints below.
Equivalently, the conditional coefficient law is Gaussian with effective
variance
\begin{equation}
\beta_j\mid\lambda_j,\tau,\zeta
\sim
\Normal(0,V_j),\qquad
V_j=\left(\zeta^{-2}+\tau^{-2}\lambda_j^{-2}\right)^{-1}
=\frac{\zeta^2\tau^2\lambda_j^2}{\zeta^2+\tau^2\lambda_j^2}.
\label{eq:supp_rhs_cond_beta}
\end{equation}
For inverse-gamma scale updates, use the half-Cauchy augmentation
\citep{MakalicSchmidt2016SimpleSampler}:
\begin{align}
\lambda_j^2\mid\nu_j&\sim\IG(1/2,1/\nu_j),&
\nu_j&\sim\IG(1/2,1),\\
\tau^2\mid\xi&\sim\IG(1/2,1/\xi),&
\xi&\sim\IG(1/2,1/\tau_0^2),
\label{eq:supp_hs_aux}\\
\zeta^2&\sim\IG(a_\zeta,b_\zeta),
\label{eq:supp_zeta_prior}
\end{align}
where the last line is omitted when \(\zeta^2\) is fixed.

\subsection{Complete-Data Posterior Targets}

The augmented posterior targets below include each AL or exAL mixture variable
once. The inverse-gamma variables in the \texttt{rhs\_ns} prior augment only
the shrinkage scales, not the likelihood.

Let \(\vect v=(v_1,\ldots,v_T)^\top\) and
\(\vect s=(s_1,\ldots,s_T)^\top\). Under exAL and ridge,
\begin{align}
p(\vect y,\vect v,\vect s,\vect\beta,\sigma,\gamma\mid\mat X)
&=
\left\{\prod_{t=1}^T
p(y_t\mid\vect\beta,\sigma,\gamma,v_t,s_t,\vect x_t)
p(v_t\mid\sigma)p(s_t)\right\}
p(\vect\beta)p(\sigma)\pi_\gamma(\gamma)\ind\{L<\gamma<U\}.
\label{eq:supp_joint_exal_ridge}
\end{align}
Under exAL and \texttt{rhs\_ns}, with
\(\Theta_{\mathrm{rhs}}=(\vect\lambda^2,\vect\nu,\tau^2,\xi,\zeta^2)\),
\begin{align}
&p(\vect y,\vect v,\vect s,\vect\beta,\sigma,\gamma,\Theta_{\mathrm{rhs}}\mid\mat X)
\nonumber\\
&\quad=
\left\{\prod_{t=1}^T
p(y_t\mid\vect\beta,\sigma,\gamma,v_t,s_t,\vect x_t)
p(v_t\mid\sigma)p(s_t)\right\}
p(\beta_0)
\prod_{j=1}^r
\left\{\Normal(\beta_j\mid0,\tau^2\lambda_j^2)
\Normal(0\mid\beta_j,\zeta^2)\right\}
\nonumber\\
&\qquad\times
\prod_{j=1}^r p(\lambda_j^2\mid\nu_j)p(\nu_j)
p(\tau^2\mid\xi)p(\xi)p(\zeta^2)
p(\sigma)\pi_\gamma(\gamma)\ind\{L<\gamma<U\}.
\label{eq:supp_joint_exal_rhs}
\end{align}
The AL joint distributions are obtained by dropping \(\vect s\) and \(\gamma\),
replacing \(A_\gamma,B_\gamma\) by \(A_0,B_0\), and using
\eqref{eq:supp_al_hierarchy}. Thus, for example, the AL target with a ridge
prior is
\begin{align}
p(\vect y,\vect v,\vect\beta,\sigma\mid\mat X)
&=
\left\{\prod_{t=1}^T
p(y_t\mid\vect\beta,\sigma,v_t,\vect x_t)p(v_t\mid\sigma)\right\}
p(\vect\beta)p(\sigma).
\label{eq:supp_joint_al_ridge}
\end{align}
The AL target with \texttt{rhs\_ns} is, analogously,
\begin{align}
&p(\vect y,\vect v,\vect\beta,\sigma,\Theta_{\mathrm{rhs}}\mid\mat X)
\nonumber\\
&\quad=
\left\{\prod_{t=1}^T
p(y_t\mid\vect\beta,\sigma,v_t,\vect x_t)p(v_t\mid\sigma)\right\}
p(\beta_0)
\prod_{j=1}^r
\left\{\Normal(\beta_j\mid0,\tau^2\lambda_j^2)
\Normal(0\mid\beta_j,\zeta^2)\right\}
\nonumber\\
&\qquad\times
\prod_{j=1}^r p(\lambda_j^2\mid\nu_j)p(\nu_j)
p(\tau^2\mid\xi)p(\xi)p(\zeta^2)p(\sigma).
\label{eq:supp_joint_al_rhs}
\end{align}
The posterior distribution is proportional to the corresponding joint density,
viewed as a function of unknown parameters and latent variables conditional on
\((\vect y,\mat X)\).

\section{MCMC Full Conditionals}
\label{sec:mcmc}

The full conditionals below apply to the Q--DESN readout after the
reservoir features are fixed. The ridge and \texttt{rhs\_ns} samplers
differ only through the coefficient prior precision and the presence of
the \texttt{rhs\_ns} scale updates. Conditional on the complete-data joint
in Section~\ref{sec:model_joint}, the Gaussian, GIG, truncated-normal,
inverse-gamma, and partially Gibbs updates below are exact; only the joint
exAL \((\sigma,\gamma)\) block needs a non-conjugate sampling step if it is
not split conditionally. The \texttt{rhs\_ns} shrinkage hierarchy is
therefore not the source of the non-conjugate posterior block under this
augmentation.

\subsection{Coefficient Block}

For exAL, define
\begin{equation}
y_t^\star=y_t-\sigma D_\gamma s_t-A_\gamma v_t,\qquad
w_t=(\sigma B_\gamma v_t)^{-1},\qquad
\mat\Omega=\diag(w_1,\ldots,w_T).
\end{equation}
Let \(\vect y^\star=(y_1^\star,\ldots,y_T^\star)^\top\). For a ridge prior,
set \(\mat P=\mat P_\beta\) and \(\vect b=\vect b_\beta\). For
\texttt{rhs\_ns}, set
\begin{equation}
\mat P=\diag\left(V_0^{-1},
\tau^{-2}\lambda_1^{-2}+\zeta^{-2},\ldots,
\tau^{-2}\lambda_r^{-2}+\zeta^{-2}\right),
\qquad
\vect b=(b_0,0,\ldots,0)^\top.
\label{eq:supp_prior_precision_mcmc}
\end{equation}
Then
\begin{align}
\mat\Sigma_{\beta\mid\cdot}
&=(\mat X^\top\mat\Omega\mat X+\mat P)^{-1},\\
\vect m_{\beta\mid\cdot}
&=\mat\Sigma_{\beta\mid\cdot}
(\mat X^\top\mat\Omega\vect y^\star+\mat P\vect b),\\
\vect\beta\mid\cdot
&\sim
\Normal(\vect m_{\beta\mid\cdot},\mat\Sigma_{\beta\mid\cdot}).
\label{eq:supp_beta_mcmc}
\end{align}
For AL, use the same update with
\[
y_t^\star=y_t-A_0v_t,\qquad
w_t=(\sigma B_0v_t)^{-1},\qquad
\mat\Omega=\diag(w_1,\ldots,w_T).
\]

\subsection{exAL Latent-Variable Blocks}

For \(v_t\), define
\[
\delta_t=y_t-\vect x_t^\top\vect\beta-\sigma D_\gamma s_t,\qquad
\chi_t=\frac{\delta_t^2}{\sigma B_\gamma},\qquad
\psi_t=\frac{A_\gamma^2}{\sigma B_\gamma}+\frac{2}{\sigma}.
\]
Then
\begin{equation}
v_t\mid\cdot\sim\GIG(1/2,\chi_t,\psi_t).
\label{eq:supp_v_mcmc}
\end{equation}
For \(s_t\), define \(\mu_{0t}=y_t-\vect x_t^\top\vect\beta-A_\gamma v_t\).
Then
\begin{align}
a_{s,t}&=1+\frac{\sigma D_\gamma^2}{B_\gamma v_t},&
m_{s,t}&=\frac{D_\gamma\mu_{0t}}{B_\gamma v_t+\sigma D_\gamma^2},\\
s_t\mid\cdot&\sim\TN(m_{s,t},a_{s,t}^{-1}).
\label{eq:supp_s_mcmc}
\end{align}

\subsection{AL Latent-Variable and Scale Blocks}

For AL,
\begin{equation}
v_t\mid\cdot
\sim
\GIG\left(
1/2,
\frac{(y_t-\vect x_t^\top\vect\beta)^2}{\sigma B_0},
\frac{A_0^2}{\sigma B_0}+\frac{2}{\sigma}
\right).
\label{eq:supp_al_v_mcmc}
\end{equation}
With \(\sigma\sim\IG(a_\sigma,b_\sigma)\),
\begin{align}
\sigma\mid\cdot
&\sim
\IG\left(
a_\sigma+\frac{3T}{2},
b_\sigma+\sum_{t=1}^T v_t
+\frac12\sum_{t=1}^T
\frac{(y_t-\vect x_t^\top\vect\beta-A_0v_t)^2}{B_0v_t}
\right).
\label{eq:supp_al_sigma_mcmc}
\end{align}

\subsection{exAL Scale and Asymmetry Blocks}

The pair \((\sigma,\gamma)\) is smooth but non-conjugate when updated jointly.
The joint log-kernel is
\begin{align}
\ell_{\sigma,\gamma}
&=
-\frac12\sum_{t=1}^T\log(\sigma B_\gamma v_t)
-\frac12\sum_{t=1}^T
\frac{\{y_t-\vect x_t^\top\vect\beta-\sigma D_\gamma s_t-A_\gamma v_t\}^2}
{\sigma B_\gamma v_t}
\nonumber\\
&\quad
-T\log\sigma-\sum_{t=1}^T\frac{v_t}{\sigma}
-(a_\sigma+1)\log\sigma-\frac{b_\sigma}{\sigma}
+\log\pi_\gamma(\gamma),
\label{eq:supp_sigmagam_joint_kernel}
\end{align}
for \(L<\gamma<U\), with value \(-\infty\) outside this interval.
This kernel can be sampled by random-walk Metropolis, slice, or a
Laplace-informed proposal on transformed coordinates. If \(\gamma\) is held
fixed while updating \(\sigma\), a partially Gibbs implementation has
\begin{align}
\sigma\mid\gamma,\cdot
&\sim
\GIG\left(
-a_\sigma-\frac{3T}{2},
2b_\sigma+2\sum_{t=1}^T v_t
+\sum_{t=1}^T\frac{(y_t-\vect x_t^\top\vect\beta-A_\gamma v_t)^2}
{B_\gamma v_t},
\sum_{t=1}^T\frac{D_\gamma^2s_t^2}{B_\gamma v_t}
\right).
\label{eq:supp_exal_sigma_mcmc}
\end{align}
Conditioning on \(\sigma\), the one-dimensional \(\gamma\) kernel is
\begin{align}
\pi(\gamma\mid\cdot)
&\propto
\pi_\gamma(\gamma)\ind\{L<\gamma<U\}
B_\gamma^{-T/2}
\exp\left[
\sum_{t=1}^T
\frac{D_\gamma s_t\{y_t-\vect x_t^\top\vect\beta-A_\gamma v_t\}}
{B_\gamma v_t}
\right.
\nonumber\\
&\qquad\left.
-\frac{1}{2\sigma}\sum_{t=1}^T
\frac{(y_t-\vect x_t^\top\vect\beta-A_\gamma v_t)^2}{B_\gamma v_t}
-\frac{\sigma}{2}\sum_{t=1}^T
\frac{D_\gamma^2s_t^2}{B_\gamma v_t}
\right].
\label{eq:supp_gamma_mcmc}
\end{align}

\subsection{Regularized-Horseshoe Shrinkage Blocks}
\label{sec:rhs_ns}

Conditioning on \(\vect\beta\), the \texttt{rhs\_ns} scale block has
inverse-gamma full conditionals:
\begin{align}
\lambda_j^2\mid\cdot
&\sim
\IG\left(1,\frac{1}{\nu_j}+\frac{\beta_j^2}{2\tau^2}\right),\\
\nu_j\mid\cdot
&\sim
\IG\left(1,1+\frac{1}{\lambda_j^2}\right),\\
\tau^2\mid\cdot
&\sim
\IG\left(\frac{r+1}{2},
\frac{1}{\xi}+\frac12\sum_{j=1}^r\frac{\beta_j^2}{\lambda_j^2}\right),\\
\xi\mid\cdot
&\sim
\IG\left(1,\frac{1}{\tau_0^2}+\frac{1}{\tau^2}\right),\\
\zeta^2\mid\cdot
&\sim
\IG\left(a_\zeta+\frac{r}{2},
b_\zeta+\frac12\sum_{j=1}^r\beta_j^2\right),
\label{eq:supp_rhs_mcmc}
\end{align}
with the final line omitted when \(\zeta^2\) is fixed. The relevant
distinction is that the Nishimura--Suchard joint construction preserves the
ordinary global-local full conditional block given \(\vect\beta\), while the
direct Piironen--Vehtari joint prior does not preserve these inverse-gamma
updates.

\begin{suppalgorithm}{Q--DESN AL and exAL MCMC}
\label{alg:qdesn_mcmc}
\emph{Input:} response vector \(\vect y\), fixed DESN design matrix \(\mat X\),
target quantile \(p_0\), likelihood choice AL or exAL, coefficient prior ridge
or \texttt{rhs\_ns}, and MCMC controls.
\begin{enumerate}[label=(\arabic*)]
\item Initialize \(\vect\beta,\sigma,\vect v\), and, for exAL,
\(\gamma,\vect s\). Initialize \texttt{rhs\_ns} scales when used.
\item For each MCMC iteration:
  \begin{enumerate}[label=(\alph*)]
  \item update \(v_t\) from \eqref{eq:supp_v_mcmc} for exAL or
  \eqref{eq:supp_al_v_mcmc} for AL;
  \item for exAL, update \(s_t\) from \eqref{eq:supp_s_mcmc};
  \item update \(\vect\beta\) from \eqref{eq:supp_beta_mcmc};
  \item for \texttt{rhs\_ns}, update
  \((\lambda_j^2,\nu_j)_{j=1}^r,\tau^2,\xi,\zeta^2\) by
  \eqref{eq:supp_rhs_mcmc};
  \item update \(\sigma\), and for exAL also update \(\gamma\), using
  \eqref{eq:supp_al_sigma_mcmc}, \eqref{eq:supp_sigmagam_joint_kernel}, or
  the partially Gibbs blocks \eqref{eq:supp_exal_sigma_mcmc}--\eqref{eq:supp_gamma_mcmc};
  \item retain draws after burn-in, with thinning only if used.
  \end{enumerate}
\end{enumerate}
\end{suppalgorithm}

\section{Variational Bayes and Laplace--Delta Updates}
\label{sec:cavi}

The VB approximation targets the same complete-data posterior as the MCMC
sampler but restricts it to a mean-field family. Gaussian, GIG,
truncated-normal, and inverse-gamma factors remain closed form, while the
non-conjugate exAL scale and asymmetry factor \((\sigma,\gamma)\) is
approximated
using Laplace--Delta on transformed coordinates
\citep{wang2013nonconjugatevb,barata2021_flex_quantile}. The update equations
therefore separate closed-form factors from Delta-method approximations to
smooth expectations involving \((\sigma,\gamma)\); the shrinkage-scale factors
remain closed form under \texttt{rhs\_ns}.

\subsection{Variational Families}

For exAL with ridge prior,
\begin{equation}
q(\vect\beta,\vect v,\vect s,\sigma,\gamma)
=q_\beta(\vect\beta)q_{\sigma,\gamma}(\sigma,\gamma)
\prod_{t=1}^Tq_{v_t}(v_t)q_{s_t}(s_t).
\label{eq:supp_vb_exal_ridge}
\end{equation}
For exAL with \texttt{rhs\_ns}, multiply \eqref{eq:supp_vb_exal_ridge} by
\[
\left\{\prod_{j=1}^rq_{\lambda_j^2}(\lambda_j^2)q_{\nu_j}(\nu_j)\right\}
q_{\tau^2}(\tau^2)q_\xi(\xi)q_{\zeta^2}(\zeta^2),
\]
omitting \(q_{\zeta^2}\) when \(\zeta^2\) is fixed. For AL, replace
\(q_{\sigma,\gamma}\) by \(q_\sigma(\sigma)\), omit \(q_{s_t}\), and set
\(\gamma=0\).

\subsection{Coefficient Factor}

Let \(\bar{\mat\Omega}=\diag(\bar w_1,\ldots,\bar w_T)\), where
\[
\bar w_t=\E\{(\sigma B_\gamma v_t)^{-1}\}
\]
for exAL and \(\bar w_t=B_0^{-1}\E(\sigma^{-1})\E(v_t^{-1})\) for AL.
For exAL define
\[
\bar m_t=
\E\left[
\frac{y_t-\sigma D_\gamma s_t-A_\gamma v_t}{\sigma B_\gamma v_t}
\right],
\qquad
\bar{\vect m}=(\bar m_1,\ldots,\bar m_T)^\top.
\]
For AL, use
\[
\bar m_t=
B_0^{-1}\left\{\E(\sigma^{-1})\E(v_t^{-1})y_t
-A_0\E(\sigma^{-1})\right\}.
\]
The prior precision is \(\bar{\mat P}=\mat P_\beta\) for ridge and
\[
\bar{\mat P}=
\diag\left(V_0^{-1},
\E(\tau^{-2})\E(\lambda_1^{-2})+\E(\zeta^{-2}),\ldots,
\E(\tau^{-2})\E(\lambda_r^{-2})+\E(\zeta^{-2})\right)
\]
for \texttt{rhs\_ns}. The prior mean vector is
\(\vect b=\vect b_\beta\) for ridge and
\(\vect b=(b_0,0,\ldots,0)^\top\) for \texttt{rhs\_ns}. Then
\begin{align}
q_\beta(\vect\beta)&=\Normal(\vect m_\beta,\mat\Sigma_\beta),\\
\mat\Sigma_\beta&=(\mat X^\top\bar{\mat\Omega}\mat X+\bar{\mat P})^{-1},\\
\vect m_\beta&=\mat\Sigma_\beta(\mat X^\top\bar{\vect m}+\bar{\mat P}\vect b).
\label{eq:supp_vb_beta}
\end{align}

\subsection{AL CAVI Updates}

Let
\[
\bar r_t=y_t-\vect x_t^\top\vect m_\beta,\qquad
\overline{r_t^2}=\bar r_t^2+\vect x_t^\top\mat\Sigma_\beta\vect x_t.
\]
For AL,
\begin{align}
q(v_t)
&=
\GIG\left(
1/2,
B_0^{-1}\E(\sigma^{-1})\overline{r_t^2},
\E(\sigma^{-1})\left(\frac{A_0^2}{B_0}+2\right)
\right),
\label{eq:supp_vb_al_v}\\
q(\sigma)
&=
\IG(a_\sigma^\star,b_\sigma^\star),\\
a_\sigma^\star
&=a_\sigma+\frac{3T}{2},\\
b_\sigma^\star
&=
b_\sigma+\sum_{t=1}^T\E(v_t)
+\frac{1}{2B_0}\sum_{t=1}^T
\left\{
\overline{r_t^2}\E(v_t^{-1})-2A_0\bar r_t+A_0^2\E(v_t)
\right\}.
\label{eq:supp_vb_al_sigma}
\end{align}

\subsection{exAL Latent-Factor Updates}

For exAL, define the smooth expectations
\[
M_1=\E\{(\sigma B_\gamma)^{-1}\},\quad
M_2=\E(D_\gamma/B_\gamma),\quad
M_3=\E(\sigma D_\gamma^2/B_\gamma),\quad
M_4=\E(A_\gamma D_\gamma/B_\gamma),
\]
and
\[
M_5=\E\{A_\gamma^2/(\sigma B_\gamma)+2/\sigma\}.
\]
These expectations are computed from the current \(q_{\sigma,\gamma}\) by
Laplace--Delta. Then
\begin{align}
q(v_t)
&=
\GIG(1/2,\bar\chi_t,\bar\psi_t),\\
\bar\chi_t
&=
M_1\overline{r_t^2}
-2M_2\bar r_t\E(s_t)
+M_3\E(s_t^2),\\
\bar\psi_t&=M_5.
\label{eq:supp_vb_exal_v}
\end{align}
The truncated-normal factor for \(s_t\) is
\begin{align}
q(s_t)&=\TN(m_{s,t}^{\mathrm{VB}},V_{s,t}^{\mathrm{VB}}),\\
V_{s,t}^{\mathrm{VB}}
&=\left\{1+M_3\E(v_t^{-1})\right\}^{-1},\\
m_{s,t}^{\mathrm{VB}}
&=
V_{s,t}^{\mathrm{VB}}
\left\{
M_2\bar r_t\E(v_t^{-1})-M_4
\right\}.
\label{eq:supp_vb_exal_s}
\end{align}

\subsection{Regularized-Horseshoe CAVI Updates}

For \(j=1,\ldots,r\), let
\(\overline{\beta_j^2}=m_{\beta,j}^2+(\mat\Sigma_\beta)_{jj}\). The
\texttt{rhs\_ns} factors are inverse-gamma:
\begin{align}
q(\lambda_j^2)
&=\IG\left(1,\ \E(\nu_j^{-1})
+\frac12\overline{\beta_j^2}\E(\tau^{-2})\right),\\
q(\nu_j)
&=\IG\left(1,\ 1+\E(\lambda_j^{-2})\right),\\
q(\tau^2)
&=\IG\left(\frac{r+1}{2},
\E(\xi^{-1})
+\frac12\sum_{j=1}^r\overline{\beta_j^2}\E(\lambda_j^{-2})\right),\\
q(\xi)
&=\IG\left(1,\ \tau_0^{-2}+\E(\tau^{-2})\right),\\
q(\zeta^2)
&=\IG\left(a_\zeta+\frac{r}{2},
b_\zeta+\frac12\sum_{j=1}^r\overline{\beta_j^2}\right).
\label{eq:supp_vb_rhs}
\end{align}
Again, the final line is omitted when \(\zeta^2\) is fixed.

\subsection{Laplace--Delta Update for \texorpdfstring{\(q_{\sigma,\gamma}\)}{q sigma gamma}}

Use unconstrained coordinates
\[
\eta_\sigma=\log\sigma,\qquad
\eta_\gamma=\logit\{(\gamma-L)/(U-L)\}.
\]
Let \(\vect\eta=(\eta_\gamma,\eta_\sigma)^\top\) and let
\((\gamma,\sigma)=T(\vect\eta)\). The variational log-target is
\[
\ell_{\mathrm{LD}}(\vect\eta)
=
\E_{q_{-(\sigma,\gamma)}}\{
\log p(\vect y,\vect v,\vect s,\vect\beta,\sigma,\gamma,\Theta_{\mathrm{rhs}}
\mid\mat X)
\}
+\log|J_T(\vect\eta)|.
\label{eq:supp_ld_logtarget}
\]
At the mode \(\hat{\vect\eta}\), with Hessian
\(\mat H=\nabla^2\ell_{\mathrm{LD}}(\hat{\vect\eta})\), set
\[
q_{\sigma,\gamma}(\vect\eta)
\approx
\Normal\{\hat{\vect\eta},(-\mat H)^{-1}\}.
\label{eq:supp_ld_gaussian}
\]
For any smooth function \(h(\sigma,\gamma)\),
\begin{equation}
\E\{h(\sigma,\gamma)\}
\approx
h\{T(\hat{\vect\eta})\}
+\frac12\tr\left[
\nabla^2_{\eta} h\{T(\hat{\vect\eta})\}(-\mat H)^{-1}
\right].
\label{eq:supp_delta}
\end{equation}
The expectations \(M_1,\ldots,M_5\), \(\E(\log\sigma)\),
\(\E(\sigma^{-1})\), \(\E(\log B_\gamma)\), and
\(\E\{\log\pi_\gamma(\gamma)\}\) are evaluated this way.

\begin{suppalgorithm}{Q--DESN Variational Bayes Updates}
\label{alg:qdesn_cavi}
\emph{Input:} response vector \(\vect y\), fixed DESN design matrix \(\mat X\),
target quantile \(p_0\), likelihood choice AL or exAL, coefficient prior ridge
or \texttt{rhs\_ns}, and convergence controls.
\begin{enumerate}[label=(\arabic*)]
\item Initialize \(q_\beta\), latent factors, scale factors, and
\texttt{rhs\_ns} factors when used.
\item Repeat until convergence:
  \begin{enumerate}[label=(\alph*)]
  \item update \(q(v_t)\) by \eqref{eq:supp_vb_al_v} for AL or
  \eqref{eq:supp_vb_exal_v} for exAL;
  \item for exAL, update \(q(s_t)\) by \eqref{eq:supp_vb_exal_s};
  \item update \(q_\beta\) by \eqref{eq:supp_vb_beta};
  \item update \texttt{rhs\_ns} factors by \eqref{eq:supp_vb_rhs} when used;
  \item update \(q(\sigma)\) for AL by \eqref{eq:supp_vb_al_sigma};
  \item update \(q_{\sigma,\gamma}\) for exAL by
  \eqref{eq:supp_ld_logtarget}--\eqref{eq:supp_ld_gaussian};
  \item recompute needed moments and monitor the ELBO.
  \end{enumerate}
\item Return variational factors and posterior summaries for the readout
quantile path \(\mu_t=\vect x_t^\top\vect\beta\).
\end{enumerate}
\end{suppalgorithm}

\section{ELBO Decomposition}
\label{sec:elbo}

The ELBO decomposition follows the complete-data joint distribution in
Section~\ref{sec:model_joint}. For exAL, the ELBO is
\begin{align}
\elbo(q)
&=
\E_q\{\log p(\vect y\mid\vect\beta,\sigma,\gamma,\vect v,\vect s,\mat X)\}
+\E_q\{\log p(\vect v\mid\sigma)\}
+\E_q\{\log p(\vect s)\}
\nonumber\\
&\quad
+\E_q\{\log p(\vect\beta,\Theta_{\mathrm{rhs}})\}
+\E_q\{\log p(\sigma)\}
+\E_q\{\log\pi_\gamma(\gamma)\}
+H(q).
\label{eq:supp_elbo_decomp}
\end{align}
For ridge, the fourth term is \(\E_q\{\log p(\vect\beta)\}\). For
\texttt{rhs\_ns}, it includes the coefficient, local-scale, global-scale,
slab, and auxiliary-variable terms in Section~\ref{subsec:rhs_elbo}.

\subsection{Likelihood and Latent Terms}

Let
\[
r_t=y_t-\vect x_t^\top\vect m_\beta,\qquad
q_t=\vect x_t^\top\mat\Sigma_\beta\vect x_t,
\]
and use expectations under the current variational factors. The exAL
likelihood contribution is
\begin{align}
\elbo_{\mathrm{like}}
&=
-\frac{T}{2}\log(2\pi)
-\frac12\sum_{t=1}^T
\left[
\E(\log\sigma)+\E(\log B_\gamma)+\E(\log v_t)
\right]
\nonumber\\
&\quad
-\frac12\sum_{t=1}^T
\E\left[
\frac{\{y_t-\vect x_t^\top\vect\beta-\sigma D_\gamma s_t-A_\gamma v_t\}^2}
{\sigma B_\gamma v_t}
\right].
\label{eq:supp_elbo_like}
\end{align}
The final expectation is exact for AL and evaluated by the same
Laplace--Delta approximations to smooth expectations used in the CAVI
updates for exAL. The latent-prior
terms are
\begin{align}
\elbo_v
&=
-T\E(\log\sigma)-\E(\sigma^{-1})\sum_{t=1}^T\E(v_t),\\
\elbo_s
&=
T\log2-\frac{T}{2}\log(2\pi)-\frac12\sum_{t=1}^T\E(s_t^2),
\label{eq:supp_elbo_latents}
\end{align}
with \(\elbo_s\) omitted for AL. The scale and asymmetry prior terms are
\begin{align}
\elbo_\sigma
&=
a_\sigma\log b_\sigma-\log\Gamma(a_\sigma)
-(a_\sigma+1)\E(\log\sigma)-b_\sigma\E(\sigma^{-1}),\\
\elbo_\gamma&=\E\{\log\pi_\gamma(\gamma)\},
\label{eq:supp_elbo_sigmagam}
\end{align}
with \(\elbo_\gamma\) omitted for AL.

\subsection{Ridge Coefficient-Prior Term}

Under the ridge prior \eqref{eq:supp_ridge_prior},
\begin{align}
\elbo_\beta^{\mathrm{ridge}}
&=
-\frac{r+1}{2}\log(2\pi)
-\frac12\log|\mat V_\beta|
\nonumber\\
&\quad
-\frac12
\left[
(\vect m_\beta-\vect b_\beta)^\top\mat P_\beta(\vect m_\beta-\vect b_\beta)
+\tr(\mat P_\beta\mat\Sigma_\beta)
\right].
\label{eq:supp_elbo_ridge}
\end{align}

\subsection{Regularized-Horseshoe Coefficient-Prior and Scale Terms}
\label{subsec:rhs_elbo}

For \(x\sim\IG(a,b)\), define
\[
m_{\log x}(a,b)=\log b-\psi(a),\qquad
m_{x^{-1}}(a,b)=a/b,\qquad
h_{\IG}(a,b)=a+\log b+\log\Gamma(a)-(a+1)\psi(a).
\]
Here \(\psi(\cdot)\) is the digamma function.
Let
\[
\ell_{\lambda j}=\E\log\lambda_j^2,\quad
r_{\lambda j}=\E(\lambda_j^{-2}),\quad
\ell_{\nu j}=\E\log\nu_j,\quad
r_{\nu j}=\E(\nu_j^{-1}),
\]
and similarly \((\ell_\tau,r_\tau)\), \((\ell_\xi,r_\xi)\), and
\((\ell_\zeta,r_\zeta)\). If \(\zeta^2\) is fixed at \(\zeta_0^2\), use
\(\ell_\zeta=\log\zeta_0^2\) and \(r_\zeta=\zeta_0^{-2}\). With
\(\overline{\beta_j^2}=m_{\beta,j}^2+(\mat\Sigma_\beta)_{jj}\),
\begin{align}
\elbo_{\beta,\mathrm{hs}}
&=
\sum_{j=1}^r
\left[
-\frac12\log(2\pi)-\frac12\ell_\tau-\frac12\ell_{\lambda j}
-\frac12\overline{\beta_j^2}r_\tau r_{\lambda j}
\right],\\
\elbo_{\beta,\mathrm{slab}}
&=
\sum_{j=1}^r
\left[
-\frac12\log(2\pi)-\frac12\ell_\zeta
-\frac12\overline{\beta_j^2}r_\zeta
\right].
\label{eq:supp_elbo_rhs_beta}
\end{align}
The half-Cauchy auxiliary-prior terms are
\begin{align}
\elbo_{\lambda\mid\nu}
&=
\sum_{j=1}^r
\left[
-\frac12\ell_{\nu j}-\log\Gamma(1/2)
-\frac32\ell_{\lambda j}-r_{\nu j}r_{\lambda j}
\right],\\
\elbo_{\nu}
&=
\sum_{j=1}^r
\left[
-\log\Gamma(1/2)-\frac32\ell_{\nu j}-r_{\nu j}
\right],\\
\elbo_{\tau\mid\xi}
&=
-\frac12\ell_\xi-\log\Gamma(1/2)-\frac32\ell_\tau-r_\xi r_\tau,\\
\elbo_{\xi}
&=
\frac12\log(\tau_0^{-2})-\log\Gamma(1/2)-\frac32\ell_\xi-\tau_0^{-2}r_\xi,\\
\elbo_{\zeta}
&=
\begin{cases}
a_\zeta\log b_\zeta-\log\Gamma(a_\zeta)
-(a_\zeta+1)\ell_\zeta-b_\zeta r_\zeta,
& \zeta^2\text{ random},\\
0, & \zeta^2\text{ fixed},
\end{cases}
\label{eq:supp_elbo_rhs_scales}
\end{align}
The unshrunk intercept prior contributes
\begin{equation}
\elbo_{\beta_0}
=
-\frac12\log(2\pi V_0)
-\frac12V_0^{-1}\{(m_{\beta,0}-b_0)^2+(\mat\Sigma_\beta)_{00}\}.
\label{eq:supp_elbo_intercept}
\end{equation}
The \texttt{rhs\_ns} contribution to the expected complete-data log density is
the sum of \eqref{eq:supp_elbo_rhs_beta}--\eqref{eq:supp_elbo_intercept}. The
entropy of the inverse-gamma scale factors is added separately through
\(H(q)\).

\subsection{Entropy Terms and Exact-versus-Approximate Status}

The Gaussian entropy is
\[
H\{q_\beta\}=\frac12\{(r+1)(1+\log2\pi)+\log|\mat\Sigma_\beta|\}.
\]
For \(x\sim\IG(a,b)\), the entropy is \(h_{\IG}(a,b)\). For
\(x\sim\GIG(\lambda,\chi,\psi)\), with \(z=(\chi\psi)^{1/2}\),
\begin{align}
H\{\GIG(\lambda,\chi,\psi)\}
&=
\log\{2K_\lambda(z)\}
-\frac{\lambda}{2}\log(\psi/\chi)
-(\lambda-1)\E(\log x)
\nonumber\\
&\quad
+\frac12\{\chi\E(x^{-1})+\psi\E(x)\}.
\label{eq:supp_entropy_gig}
\end{align}
For \(s\sim\TN(m,V)\), let \(a=V^{1/2}\), \(z=m/a\), and
\(\Lambda(z)=\phi(z)/\Phi(z)\). Then
\[
H\{\TN(m,V)\}
=
\frac12\log(2\pi V)+\log\Phi(z)+\frac12\{1-z\Lambda(z)\}.
\]
For the Laplace--Delta block, if
\(q_\eta=\Normal(\hat{\vect\eta},\mat\Sigma_\eta)\), then
\[
H\{q_{\sigma,\gamma}\}
\approx
\frac12\{2(1+\log2\pi)+\log|\mat\Sigma_\eta|\}
+\E_{q_\eta}\{\log|J_T(\vect\eta)|\}.
\]
All Gaussian, inverse-gamma, GIG, and truncated-normal terms are exact up to
standard special-function evaluation. Terms involving smooth nonlinear
functions of \((\sigma,\gamma)\) are approximate under the
Laplace--Delta expansion.

\section{Joint exAL--QVP--Q--DESN--RHS Extension}
\label{sec:joint_qdesn_supp}

This section gives the derivation layer for the joint quantile-vector readout
introduced in the main article. The reservoir construction is unchanged. The
only new object is a prior over the quantile-indexed readout slope path,
conditional on the fixed non-intercept DESN design
\(\mat Z=[\tilde{\vect x}_1^\top;\ldots;\tilde{\vect x}_T^\top]\in\R^{T\times r}\).
The intercepts are kept separate from the high-dimensional slope shrinkage.

\subsection{Stacked exAL Working Likelihood}

Let \(0<p_1<\cdots<p_Q<1\) be the quantile grid. At level \(p_q\),
\[
Q_{p_q}(y_t\mid\mathcal F_t)=
\alpha_q+\tilde{\vect x}_t^\top\vect\beta_q,
\qquad
\alpha_q\in\R,\quad \vect\beta_q\in\R^r .
\]
This notation contains the single-level RHS Q--DESN as the \(Q=1\) convention
\(\alpha_1=\alpha\), \(\vect\beta_1=\vect\beta\), with no learned slope
baseline. Equivalently, for the single-quantile limit set
\(\vect\beta_{\mathrm{base}}\equiv\zeros{r}\) and assign the RHS prior
directly to \(\Delta_1=\vect\beta_1\). The baseline in the multi-quantile
derivation below is therefore a slope-path device, not an intercept term. In
this limit, the baseline RHS hierarchy and its variational factors are simply
absent; the innovation RHS hierarchy is the ordinary single-level RHS prior.
The exAL augmented likelihood is a quantile-indexed copy of
\eqref{eq:supp_exal_hierarchy}:
\begin{align}
v_{q,t}\mid\sigma_q&\sim\Exp(\text{rate}=1/\sigma_q),&
s_{q,t}&\sim\TN(0,1),\\
y_t\mid\alpha_q,\vect\beta_q,\sigma_q,\gamma_q,v_{q,t},s_{q,t}
&\sim
\Normal\left(
\alpha_q+\tilde{\vect x}_t^\top\vect\beta_q
+\sigma_qD_qs_{q,t}+A_qv_{q,t},\,
\sigma_qB_qv_{q,t}
\right),
\label{eq:supp_joint_exal_hierarchy}
\end{align}
where \(A_q=A_{p_q}(\gamma_q)\), \(B_q=B_{p_q}(\gamma_q)\), and
\(D_q=\lambda_{p_q}(\gamma_q)\). The AL special case fixes the AL constants
and omits \(s_{q,t}\) and \(\gamma_q\).

Define
\[
\vect y_{\mathrm{st}}=\ones{Q}\otimes\vect y,\qquad
\mat Z_{\mathrm{st}}=\Id{Q}\otimes\mat Z,\qquad
\vect\beta_{\mathrm{st}}=(\vect\beta_1^\top,\ldots,\vect\beta_Q^\top)^\top,
\]
and
\[
\vect\alpha_{\mathrm{st}}
=
(\alpha_1\ones{T}^\top,\ldots,\alpha_Q\ones{T}^\top)^\top .
\]
Let
\[
d_{q,t}=\sigma_qD_qs_{q,t}+A_qv_{q,t},\qquad
\omega_{q,t}=\sigma_qB_qv_{q,t},
\]
and stack these into \(\vect d\) and
\(\mat\Omega=\diag(\omega_{q,t})\). Conditional on the latent variables and
likelihood parameters,
\begin{equation}
\vect y_{\mathrm{st}}\mid\vect\beta_{\mathrm{st}},\vect\alpha,\vect d,\mat\Omega
\sim
\Normal(
\vect\alpha_{\mathrm{st}}+\mat Z_{\mathrm{st}}\vect\beta_{\mathrm{st}}+\vect d,\,
\mat\Omega).
\label{eq:supp_joint_stacked_likelihood}
\end{equation}
If a composite-likelihood learning rate is used, the Gaussian observation term
in \eqref{eq:supp_joint_stacked_likelihood} is multiplied by
\(\kappa>0\). The default is \(\kappa=1\); \(\kappa=1/Q\) is a conservative
sensitivity setting.

\subsection{QVP--RHS Prior}
\label{subsec:supp_joint_qvp_prior}

For a genuine multi-quantile grid, let
\(\vect\beta_{\mathrm{base}}\in\R^r\) denote a baseline readout. Adjacent
innovations are
\[
\Delta_1=\vect\beta_1-\vect\beta_{\mathrm{base}},\qquad
\Delta_q=\vect\beta_q-\vect\beta_{q-1},\quad q=2,\ldots,Q.
\]
The baseline slope and adjacent innovations use separate RHS hierarchies:
\begin{align}
\beta_{\mathrm{base},j}\mid\lambda^0_j,\tau^0,\zeta_0
&\sim\Normal(0,V^0_j),&
V^0_j
&=
\frac{\zeta_0^2(\tau^0)^2(\lambda^0_j)^2}
{\zeta_0^2+(\tau^0)^2(\lambda^0_j)^2},
\label{eq:supp_joint_baseline_var}\\
\Delta_{q,j}\mid\lambda^\Delta_{q,j},\tau^\Delta_q,\zeta_\Delta
&\sim\Normal(0,V^\Delta_{q,j}),&
V^\Delta_{q,j}
&=
\frac{\zeta_\Delta^2(\tau^\Delta_q)^2(\lambda^\Delta_{q,j})^2}
{\zeta_\Delta^2+(\tau^\Delta_q)^2(\lambda^\Delta_{q,j})^2}.
\label{eq:supp_joint_innovation_var}
\end{align}
The local and global half-Cauchy scales use the inverse-gamma augmentation in
\eqref{eq:supp_hs_aux}; it is applied once to the baseline coefficients and
separately to each adjacent innovation block.

Let \(\mat H\in\R^{Qr\times Qr}\) be the block first-difference matrix
satisfying
\[
\mat H\vect\beta_{\mathrm{st}}=
(\vect\beta_1^\top,
(\vect\beta_2-\vect\beta_1)^\top,\ldots,
(\vect\beta_Q-\vect\beta_{Q-1})^\top)^\top .
\]
Set
\[
\tilde{\vect\beta}_{\mathrm{base}}=
(\vect\beta_{\mathrm{base}}^\top,\zeros{r}^\top,\ldots,\zeros{r}^\top)^\top,
\qquad
\mat D_\Delta=\diag(V^\Delta_{q,j}:q=1,\ldots,Q,\ j=1,\ldots,r).
\]
Then the conditional QVP prior can be written as
\begin{equation}
p(\vect\beta_{\mathrm{st}}\mid\vect\beta_{\mathrm{base}},\mat D_\Delta)
\propto
|\mat D_\Delta|^{-1/2}
\exp\left[
-\frac12
(\mat H\vect\beta_{\mathrm{st}}-\tilde{\vect\beta}_{\mathrm{base}})^\top
\mat D_\Delta^{-1}
(\mat H\vect\beta_{\mathrm{st}}-\tilde{\vect\beta}_{\mathrm{base}})
\right].
\label{eq:supp_joint_qvp_prior}
\end{equation}
This representation is sparse and block banded. It is the computational form
used by both the MCMC and VB updates.

\subsection{Complete-Data Posterior}

Let \(\Theta_{\Delta}\) collect all local, global, auxiliary, and slab scales
for the innovation RHS hierarchy, and let \(\Theta_0\) collect the
corresponding baseline RHS scales. Up to constants not depending on unknowns,
the complete-data posterior for the multi-quantile case is
\begin{align}
&p(\vect\beta_{\mathrm{st}},\vect\beta_{\mathrm{base}},\vect\alpha,
\{\sigma_q,\gamma_q\}_{q=1}^Q,
\{v_{q,t},s_{q,t}\}_{q,t},
\Theta_0,\Theta_\Delta\mid\vect y,\mat Z)
\nonumber\\
&\quad\propto
\exp\left[
-\frac{\kappa}{2}
(\vect y_{\mathrm{st}}-\vect\alpha_{\mathrm{st}}
-\mat Z_{\mathrm{st}}\vect\beta_{\mathrm{st}}-\vect d)^\top
\mat\Omega^{-1}
(\vect y_{\mathrm{st}}-\vect\alpha_{\mathrm{st}}
-\mat Z_{\mathrm{st}}\vect\beta_{\mathrm{st}}-\vect d)
\right]
\nonumber\\
&\qquad\times
\prod_{q=1}^Q\prod_{t=1}^T
\left\{
(\omega_{q,t})^{-\kappa/2}
p(v_{q,t}\mid\sigma_q)p(s_{q,t})
\right\}
\prod_{q=1}^Q p(\sigma_q)\pi_{\gamma_q}(\gamma_q)\ind\{L_q<\gamma_q<U_q\}
\nonumber\\
&\qquad\times
p(\vect\alpha)\,
p(\vect\beta_{\mathrm{base}}\mid\Theta_0)p(\Theta_0)\,
p(\vect\beta_{\mathrm{st}}\mid\vect\beta_{\mathrm{base}},\mat D_\Delta)p(\Theta_\Delta).
\label{eq:supp_joint_complete_posterior}
\end{align}
The factor \(\kappa\) is attached to the Gaussian observation contribution.
The latent-variable densities remain the exAL augmentation densities; this
keeps \(\kappa=1\) equal to the ordinary complete-data posterior and makes
other choices interpretable as tempered observation information.
For \(Q=1\), remove
\(\vect\beta_{\mathrm{base}}\), \(\Theta_0\), and the baseline-prior factor
from \eqref{eq:supp_joint_complete_posterior}, set
\(\Delta_1=\vect\beta_1\), and use the innovation RHS hierarchy as the
ordinary single-level RHS hierarchy.

\subsection{Gaussian Blocks}

Let
\(\vect r_\beta=\vect y_{\mathrm{st}}-\vect\alpha_{\mathrm{st}}-\vect d\).
Combining \eqref{eq:supp_joint_stacked_likelihood} and
\eqref{eq:supp_joint_qvp_prior} gives
\begin{align}
\mat K_\beta
&=
\kappa\mat Z_{\mathrm{st}}^\top\mat\Omega^{-1}\mat Z_{\mathrm{st}}
+\mat H^\top\mat D_\Delta^{-1}\mat H,\\
\vect m_\beta
&=
\mat K_\beta^{-1}
\left\{
\kappa\mat Z_{\mathrm{st}}^\top\mat\Omega^{-1}\vect r_\beta
+\mat H^\top\mat D_\Delta^{-1}\tilde{\vect\beta}_{\mathrm{base}}
\right\}.
\label{eq:supp_joint_beta_conditional}
\end{align}
Thus
\[
\vect\beta_{\mathrm{st}}\mid\cdots\sim\Normal(\vect m_\beta,\mat K_\beta^{-1}).
\]
In implementation, \(\mat K_\beta\) should be treated as a sparse precision
matrix; the dense inverse is not needed for MCMC sampling.

The baseline coefficients are conditionally independent by feature under the
diagonal RHS representation. For \(j=1,\ldots,r\),
\begin{align}
K_{0,j}&=(V^0_j)^{-1}+(V^\Delta_{1,j})^{-1},\\
m_{0,j}&=K_{0,j}^{-1}(V^\Delta_{1,j})^{-1}\beta_{1,j},
\end{align}
so
\[
\beta_{\mathrm{base},j}\mid\cdots\sim\Normal(m_{0,j},K_{0,j}^{-1}).
\]
This update uses the fact that only the first innovation
\(\Delta_1=\vect\beta_1-\vect\beta_{\mathrm{base}}\) contains \(\vect\beta_{\mathrm{base}}\).

For unconstrained intercepts with independent priors
\(\alpha_q\sim\Normal(a_{\alpha,q},V_{\alpha,q})\), define
\[
e_{q,t}^{\alpha}=y_t-d_{q,t}-\tilde{\vect x}_t^\top\vect\beta_q .
\]
Then
\begin{align}
K_{\alpha,q}
&=V_{\alpha,q}^{-1}+\kappa\sum_{t=1}^T\omega_{q,t}^{-1},\\
m_{\alpha,q}
&=K_{\alpha,q}^{-1}
\left\{
V_{\alpha,q}^{-1}a_{\alpha,q}
+\kappa\sum_{t=1}^T\omega_{q,t}^{-1}e_{q,t}^{\alpha}
\right\},
\end{align}
and \(\alpha_q\mid\cdots\sim\Normal(m_{\alpha,q},K_{\alpha,q}^{-1})\). For a
monotone-gap version, write
\(\alpha_q=\alpha_1+\sum_{\ell=2}^q\exp(\eta_\ell)\) for
\(q\ge2\). The conditional log density for
\((\alpha_1,\eta_2,\ldots,\eta_Q)\) is the Gaussian likelihood term above plus
the chosen priors on the transformed coordinates and, if the prior is placed on
the positive gaps rather than on \(\eta\), the Jacobian
\(\sum_{\ell=2}^Q\eta_\ell\). This block is then updated by a
low-dimensional Metropolis--Hastings, slice, or Laplace step.

\subsection{exAL Latent-Variable and Scale Blocks}

Conditional on the joint readout, each exAL latent update is a
quantile-indexed copy of the single-level update with the Gaussian likelihood
scaled by \(\kappa\). Define
\[
e_{q,t}^{v}
=y_t-\alpha_q-\tilde{\vect x}_t^\top\vect\beta_q-\sigma_qD_qs_{q,t}.
\]
Then
\[
p(v_{q,t}\mid\cdots)
\propto
v_{q,t}^{-\kappa/2}
\exp\left[
-\frac12\left\{
\chi_{q,t}^{v}v_{q,t}^{-1}
+\psi_{q,t}^{v}v_{q,t}
\right\}
\right],
\]
which is a GIG density with
\[
\chi_{q,t}^{v}=\frac{\kappa(e_{q,t}^{v})^2}{\sigma_qB_q},
\qquad
\psi_{q,t}^{v}=\frac{\kappa A_q^2}{\sigma_qB_q}+\frac{2}{\sigma_q}.
\]
The GIG index is \(1-\kappa/2\). Terms that do not depend on \(v_{q,t}\) have
been omitted from the kernel.

For the positive-shift latent variable, define
\[
e_{q,t}^{s}
=y_t-\alpha_q-\tilde{\vect x}_t^\top\vect\beta_q-A_qv_{q,t}.
\]
The conditional distribution is positive-truncated Normal with precision and
mean
\[
K_{s,q,t}=1+\frac{\kappa\sigma_qD_q^2}{B_qv_{q,t}},
\qquad
m_{s,q,t}=K_{s,q,t}^{-1}
\frac{\kappa D_qe_{q,t}^{s}}{B_qv_{q,t}},
\]
that is,
\[
s_{q,t}\mid\cdots\sim\TN(m_{s,q,t},K_{s,q,t}^{-1}).
\]
The AL case omits this update and uses the AL constants in the \(v_{q,t}\)
kernel.

For each quantile level, the non-conjugate scale-asymmetry block has log
kernel
\begin{align}
\ell_q(\sigma_q,\gamma_q)
&=
-\frac{\kappa}{2}\sum_{t=1}^T
\left[
\log\{\sigma_qB_qv_{q,t}\}
+
\frac{
\{y_t-\alpha_q-\tilde{\vect x}_t^\top\vect\beta_q
-\sigma_qD_qs_{q,t}-A_qv_{q,t}\}^2}
{\sigma_qB_qv_{q,t}}
\right]
\nonumber\\
&\quad
-T\log\sigma_q-\sigma_q^{-1}\sum_{t=1}^Tv_{q,t}
+\log p(\sigma_q)+\log\pi_{\gamma_q}(\gamma_q).
\label{eq:supp_joint_sigma_gamma_kernel}
\end{align}
This kernel is evaluated on transformed coordinates
\(\eta_{q,1}=\log\sigma_q\) and
\(\eta_{q,2}=\log\{(\gamma_q-L_q)/(U_q-\gamma_q)\}\), adding the
transformation Jacobian for MCMC proposals or the Laplace--Delta approximation.

\subsection{RHS Scale Updates}

The RHS scale updates are copies of the single-level
Nishimura--Suchard product updates. For a baseline coefficient,
\[
\lambda_{0,j}^2\mid\cdots
\sim
\IG\left(
1,\,
\frac{1}{\nu_{0,j}}+\frac{\beta_{\mathrm{base},j}^2}{2(\tau^0)^2}
\right),
\qquad
\nu_{0,j}\mid\cdots
\sim
\IG\left(1,\ 1+\frac{1}{\lambda_{0,j}^2}\right),
\]
and
\[
(\tau^0)^2\mid\cdots
\sim
\IG\left(
\frac{r+1}{2},\,
\frac{1}{\xi_0}
+\frac12\sum_{j=1}^r\frac{\beta_{\mathrm{base},j}^2}{(\lambda^0_j)^2}
\right),
\qquad
\xi_0\mid\cdots
\sim
\IG\left(1,\frac{1}{(\tau^0)^2}+\frac{1}{(\tau_0^0)^2}\right).
\]
For each innovation block \(q\), replace \(\beta_{\mathrm{base},j}\),
\(\lambda^0_j\), and \(\tau^0\) by \(\Delta_{q,j}\),
\(\lambda^\Delta_{q,j}\), and \(\tau^\Delta_q\). If the slab scale is random
and shared across innovation blocks, then
\[
\zeta_\Delta^2\mid\cdots
\sim
\IG\left(
a_{\zeta_\Delta}+\frac{Qr}{2},\,
b_{\zeta_\Delta}
+\frac12\sum_{q=1}^Q\sum_{j=1}^r\Delta_{q,j}^2
\right),
\]
with the analogous \(r/2\) update for a random baseline slab scale
\(\zeta_0^2\). If slab scales are fixed, these updates are omitted.

\subsection{MCMC and VB--LD Algorithms}

\begin{suppalgorithm}{Joint exAL--QVP--Q--DESN--RHS MCMC}
\label{alg:joint_qdesn_mcmc}
\emph{Input:} fixed non-intercept design \(\mat Z\), response \(\vect y\),
quantile grid \(p_1<\cdots<p_Q\), priors, learning rate \(\kappa\), and MCMC
controls. For \(Q=1\), use the collapsed convention
\(\vect\beta_{\mathrm{base}}\equiv\zeros{r}\) and
\(\Delta_1=\vect\beta_1\), and skip every baseline update below.
\begin{enumerate}[label=(\arabic*)]
\item Initialize \(\vect\alpha\), \(\vect\beta_{\mathrm{st}}\), likelihood
parameters, exAL latent variables, and innovation RHS scales. If \(Q>1\),
also initialize \(\vect\beta_{\mathrm{base}}\) and baseline RHS scales.
\item Repeat:
  \begin{enumerate}[label=(\alph*)]
  \item update \(v_{q,t}\) and \(s_{q,t}\) for all \(q,t\);
  \item update \(\vect\beta_{\mathrm{st}}\) from the sparse Gaussian precision
  \eqref{eq:supp_joint_beta_conditional};
  \item if \(Q>1\), update \(\vect\beta_{\mathrm{base}}\) featurewise;
  \item update \(\vect\alpha\) by the Gaussian block or by the monotone-gap
  transformed block;
  \item update innovation RHS scales, and if \(Q>1\) update baseline RHS
  scales;
  \item update each \((\sigma_q,\gamma_q)\) block using
  \eqref{eq:supp_joint_sigma_gamma_kernel} on transformed coordinates;
  \item store draws and crossing diagnostics after burn-in.
  \end{enumerate}
\end{enumerate}
\end{suppalgorithm}

For VB--LD, a mean-field factorization matching the MCMC blocks is
\begin{align}
q
&=
q_\beta(\vect\beta_{\mathrm{st}})q_{\beta_{\mathrm{base}}}(\vect\beta_{\mathrm{base}})q_\alpha(\vect\alpha)
\prod_{q=1}^Q q_{\sigma_q,\gamma_q}(\sigma_q,\gamma_q)
\prod_{q=1}^Q\prod_{t=1}^T q_{v_{q,t}}(v_{q,t})q_{s_{q,t}}(s_{q,t})
\nonumber\\
&\quad\times
q(\Theta_0)q(\Theta_\Delta),
\label{eq:supp_joint_vb_factorization}
\end{align}
with \(q_{s_{q,t}}\) omitted under AL. The factors
\(q_\beta\), \(q_{\beta_{\mathrm{base}}}\), and \(q_\alpha\) are Gaussian; the \(v\)
factors are GIG; the \(s\) factors are positive-truncated Normal; and the RHS
scale factors are inverse-gamma. The \(q_{\sigma_q,\gamma_q}\) factors use
Laplace--Delta approximations on the transformed coordinates above. For
\(Q=1\), remove \(q_{\beta_{\mathrm{base}}}\) and \(q(\Theta_0)\), set
\(\Delta_1=\vect\beta_1\), and read \(q_\beta\) as the ordinary
single-level slope factor.

The Gaussian variational precision for \(q_\beta\) is obtained from
\eqref{eq:supp_joint_beta_conditional} by replacing
\(\mat\Omega^{-1}\), \(\mat D_\Delta^{-1}\), \(\vect d\), and
\(\tilde{\vect\beta}_{\mathrm{base}}\) with the corresponding variational expectations.
Only selected covariance elements are needed for the RHS scale expectations and
ELBO; a full dense inverse of the stacked precision is not required when sparse
selected-inverse or diagonal-extraction routines are available.

The joint ELBO is
\begin{align}
\elbo_{\mathrm{joint}}(q)
&=
\kappa\E_q\{\log p(\vect y_{\mathrm{st}}\mid
\vect\beta_{\mathrm{st}},\vect\alpha,\vect d,\mat\Omega)\}
+\sum_{q,t}\E_q\{\log p(v_{q,t}\mid\sigma_q)\}
+\sum_{q,t}\E_q\{\log p(s_{q,t})\}
\nonumber\\
&\quad
+\sum_q\E_q\{\log p(\sigma_q)+\log\pi_{\gamma_q}(\gamma_q)\}
+\E_q\{\log p(\vect\alpha)\}
+\E_q\{\log p(\vect\beta_{\mathrm{st}}\mid\vect\beta_{\mathrm{base}},\mat D_\Delta)\}
\nonumber\\
&\quad
+\E_q\{\log p(\vect\beta_{\mathrm{base}},\Theta_0)\}
+\E_q\{\log p(\Theta_\Delta)\}
+H(q).
\label{eq:supp_joint_elbo}
\end{align}
The Gaussian, GIG, truncated-normal, and inverse-gamma terms are exact under
the variational factorization. Expectations involving nonlinear functions of
\((\sigma_q,\gamma_q)\) are evaluated with the same Laplace--Delta expansion
used in the single-level VB--LD algorithm.

\subsection{Crossing Diagnostics}

For adjacent levels \(q-1\) and \(q\), define the fitted gap at time \(t\) by
\[
g_{q,t}=
(\alpha_q-\alpha_{q-1})
+\tilde{\vect x}_t^\top(\vect\beta_q-\vect\beta_{q-1}).
\]
The raw crossing count is
\[
C_{\mathrm{raw}}=\sum_{q=2}^Q\sum_{t=1}^T\ind\{g_{q,t}<0\}.
\]
If reported quantile curves are projected or rearranged, the corresponding
reported crossing count should be computed after that declared monotone
contract. The maximum and mean absolute projection adjustments are useful
diagnostics because a zero reported crossing count alone does not show whether
the joint prior produced coherent raw paths or whether the reporting contract
made large repairs. On a bounded feature domain
\(\tilde{\vect x}_t\in[-1,1]^r\), the sufficient condition
\[
\alpha_q-\alpha_{q-1}\ge
\|\vect\beta_q-\vect\beta_{q-1}\|_1
\]
is a diagnostic reference for soft non-crossing behavior; it is not imposed by
the QVP--RHS prior unless a separate hard-constraint step is added.

\section{Posterior Prediction and Quantile Synthesis}
\label{sec:forecast_synthesis}

Forecasting propagates posterior or variational draws through the fixed
reservoir, whereas multi-quantile synthesis is a
post-processing device for independently fitted quantile levels rather than a
joint multi-quantile posterior model.

For a single-level fit at target \(p_0\), write a posterior draw or a draw from
the variational approximation as
\(\theta_{p_0}^{(m)}
=(\vect\beta_{p_0}^{(m)},\sigma_{p_0}^{(m)},\gamma_{p_0}^{(m)})\). Forecasting
proceeds by propagating the deterministic reservoir state forward from the
forecast origin. At horizon \(h\), the current lag vector is built from
observed values for one-step forecasts and from previously simulated values for
free-running multi-step forecasts. Conditional on the draw and the resulting
\(p_0\)-indexed design row, the working-likelihood draw is
\[
y_{T+h}^{(m,p_0)}
\sim
\exAL_{p_0}
\left(\vect x_{T+h}^{(m,p_0)\top}\vect\beta_{p_0}^{(m)},
\sigma_{p_0}^{(m)},\gamma_{p_0}^{(m)}\right),
\]
with the \(s\)-latent and \(\gamma\) block omitted for AL. Repeating this over
posterior or variational draws produces working-likelihood forecast draws,
model-based forecast bands, and \(p_0\)-level quantile summaries conditional on
the fixed reservoir and working likelihood. To report several independently
fitted levels, the same recursion is run separately for each fitted \(p_k\).

For a joint quantile-vector readout, a posterior or variational draw returns a
raw quantile vector over \(p_1,\ldots,p_Q\). It does not produce one replicated
future response for each quantile level. If recursive lag inputs require a
future response path, that path must be generated by a declared propagation rule.
For draw \(m\) at horizon \(h\), write
\[
\mu_{q,T+h}^{(m)}
=\alpha_q^{(m)}
+\tilde{\vect x}_{T+h}^{(m)\top}\vect\beta_q^{(m)},
\qquad q=1,\ldots,Q,
\]
and let
\(\vect q_{T+h}^{(m)}
=(\mu_{1,T+h}^{(m)},\ldots,\mu_{Q,T+h}^{(m)})^\top\). If exact monotone output
is required, apply the declared reporting contract
\(\vect q_{T+h}^{*,(m)}=\mathcal M(\vect q_{T+h}^{(m)})\). Piecewise-linear
interpolation across
\(\{(p_q,q_{q,T+h}^{*,(m)})\}_{q=1}^Q\), with the conservative endpoint rule
below, defines \(Q_{T+h}^{*,(m)}(u)\). A scalar synthesized response draw is
then
\[
u_{T+h}^{(m,r)}\sim\Unif(0,1),
\qquad
y_{T+h}^{(m,r)}
=Q_{T+h}^{*,(m)}\!\left(u_{T+h}^{(m,r)}\right).
\]
This is inverse-CDF sampling from the posterior-sampled quantile function. It is
not obtained by mixing the quantile-specific working likelihoods, normalizing
their product, or drawing separately from their AL/exAL factors.

When separate models are fit at ordered quantile levels
\(p_1<\cdots<p_K\), multi-quantile synthesis can form a monotone predictive
quantile curve. This generic post-processing device is used for independently
fitted quantile models and is not part of the Q--DESN
posterior target. At each time point, collect fitted target-quantile summaries
\(m_{1,t},\ldots,m_{K,t}\), such as posterior means or medians of the readout
locations from the \(K\) independently fitted models. The synthesis uses this
target grid rather than off-target quantiles from the working likelihoods. If
necessary, replace the grid by the isotonic projection
\(m_{1,t}^\star\le\cdots\le m_{K,t}^\star\)
\citep{BarlowBrunk1972IsotonicDual}. Linear
interpolation across \((p_k,m_{k,t}^\star)\), with conservative endpoint
extrapolation, gives the estimated quantile function. If posterior-draw-level
curves or smoother interpolants introduce residual non-monotonicity on a dense
grid, a final monotone rearrangement can be applied \citep{chernozhukov2010}.
The finite grid does not identify tail shape outside \([p_1,p_K]\); parametric
tail extension should therefore be treated as a separate modeling step rather
than an implication of the composite working likelihood.

\clearpage
\section{Quantile-Free Interval Estimation: RQR--DESN}
\label{sec:rqr_desn_supp}

The relaxed quantile regression DESN (RQR--DESN) extension changes the readout
target while keeping the DESN feature map fixed. For a target coverage level
\(c\in(0,1)\), the model uses two exchangeable root readouts
\[
\eta_{1t}=\vect x_t^\top\vect\beta_1,\qquad
\eta_{2t}=\vect x_t^\top\vect\beta_2,
\]
and reports the ordered interval endpoints
\[
L_t=\min(\eta_{1t},\eta_{2t}),\qquad
U_t=\max(\eta_{1t},\eta_{2t}).
\]
The root labels are exchangeable; lower and upper labels are applied only
after ordering the two root ordinates. The coverage level \(c\) is not a
quantile index. It indexes a direct interval target, following the relaxed
quantile regression loss of \citet{PouplinEtAl2024RQR}. With
\[
\rho_c(u)=u\{c-\ind(u<0)\},\qquad
e_t=(y_t-\eta_{1t})(y_t-\eta_{2t}),
\]
the sign of \(e_t\) records whether the observation is inside the interval:
\(e_t<0\) when \(L_t<y_t<U_t\) and \(e_t>0\) outside the interval. The
empirical RQR loss is
\[
\mathcal L_{T,c}(\vect\beta_1,\vect\beta_2)
=
\sum_{t=1}^T\rho_c(e_t),
\]
and the generalized posterior is
\[
\Pi_{\mathrm R}(d\vect\beta_1,d\vect\beta_2\mid \vect y,\mat X)
\propto
\pi_\beta(d\vect\beta_1)\pi_\beta(d\vect\beta_2)
\exp\left\{-\omega_{\mathrm R}\mathcal L_{T,c}(\vect\beta_1,\vect\beta_2)\right\},
\]
where \(\omega_{\mathrm R}>0\) is a learning rate. This is a
generalized-Bayes update \citep{BissiriHolmesWalker2016GeneralBayes}, not a
normalized response likelihood. Consequently, RQR--DESN produces posterior
summaries for interval functionals \((L_t,U_t)\), but it does not provide
posterior predictive draws of \(y_t\).

The population target has a useful interpretation under an unrestricted
interior solution with positive interval width. Writing \(L(x)\) and \(U(x)\)
for the endpoint functions, the first-order conditions for the expected RQR
loss imply
\[
\Pr\{L(x)<Y<U(x)\mid X=x\}=c
\]
and
\[
\E\!\left[
Y\ind\{L(x)<Y<U(x)\}\mid X=x
\right]
=c\,\E(Y\mid X=x).
\]
Thus the ideal target is a coverage-plus-mean-balance interval. The finite
fixed-feature readout allows asymmetric interval estimation, with empirical
coverage and generalized-posterior calibration checked by the validation
protocol rather than assumed from the loss alone.

\subsection{Pseudo-AL Augmentation}
\label{subsec:rqr_desn_aug}

The implementation uses a pseudo-AL representation of the RQR residual product.
Let \(r_t=y_t^2\) and
\[
m_t^{\mathrm R}
=y_t(\eta_{1t}+\eta_{2t})-\eta_{1t}\eta_{2t},
\qquad
r_t-m_t^{\mathrm R}=e_t.
\]
Set
\[
\sigma_{\mathrm R}=\omega_{\mathrm R}^{-1},\qquad
\xi_c=\frac{1-2c}{c(1-c)},\qquad
\phi_c=\frac{2}{c(1-c)}.
\]
The exponentiated loss contribution satisfies the exact identity
\[
\exp\{-\omega_{\mathrm R}\rho_c(e_t)\}
=
\frac{\sigma_{\mathrm R}}{c(1-c)}
\int_0^\infty
\Normal\!\left(
r_t\mid m_t^{\mathrm R}+\xi_c v,\,
\phi_c\sigma_{\mathrm R}v
\right)
\frac{1}{\sigma_{\mathrm R}}\exp(-v/\sigma_{\mathrm R})\,dv .
\]
Equivalently, conditional on \(v_t>0\), the pseudo-observation
representation is
\[
r_t\mid \vect\beta_1,\vect\beta_2,v_t
\sim
\Normal\{m_t^{\mathrm R}+\xi_c v_t,\ \phi_c\sigma_{\mathrm R} v_t\},
\qquad
v_t\sim \Exp(\text{rate}=1/\sigma_{\mathrm R}).
\]
The augmentation is therefore a computational device for the interval loss,
not a distributional model for the original response. The observed response
appears in both \(r_t\) and \(m_t^{\mathrm R}\), which is why the Gaussian
display is not a generative observation equation for \(y_t\).

With the \(\GIG(p,a,b)\) convention proportional to
\(x^{p-1}\exp\{-(ax+b/x)/2\}\), the latent-scale update is
\[
v_t\mid\cdot\sim
\GIG\left\{
\frac12,\
\frac{1}{2\sigma_{\mathrm R} c(1-c)},\
\frac{c(1-c)e_t^2}{2\sigma_{\mathrm R}}
\right\}.
\]
Numerically, the Gaussian coefficient blocks should be formed with stable
symmetric solves, such as Cholesky factorization with the same precision guards
used for the Q--DESN readout. If \(e_t\) is zero to numerical precision, the
GIG update must be evaluated using a sampler that supports the limiting case or
with a documented positive residual-product floor.

\subsection{Coefficient Blocks}
\label{subsec:rqr_desn_blocks}

Let \(W_v=\diag\{(\phi_c\sigma_{\mathrm R} v_t)^{-1}:t=1,\ldots,T\}\).
Conditional on the second root, the pseudo-observation equation for
\(\vect\beta_1\) is linear. Define
\[
\mat A_1=\diag(\vect y-\vect\eta_2)\mat X,\qquad
\vect z_1=\vect y^{\,2}-\vect y\circ\vect\eta_2-\xi_c\vect v.
\]
If the prior for \(\vect\beta_1\) has Gaussian canonical parameters
\((\mat Q_{01},\vect h_{01})\), then
\[
\mat Q_1=\mat Q_{01}+\mat A_1^\top W_v\mat A_1,\qquad
\vect h_1=\vect h_{01}+\mat A_1^\top W_v\vect z_1,
\]
and
\[
\vect\beta_1\mid\cdot\sim
\Normal(\mat Q_1^{-1}\vect h_1,\mat Q_1^{-1}).
\]
The update for \(\vect\beta_2\) is identical after exchanging the root labels:
\[
\mat A_2=\diag(\vect y-\vect\eta_1)\mat X,\qquad
\vect z_2=\vect y^{\,2}-\vect y\circ\vect\eta_1-\xi_c\vect v.
\]
Ridge and \texttt{rhs\_ns} priors enter through the same canonical Gaussian
blocks used by the Q--DESN readout, applied separately to the two root
coefficient vectors.

\begin{suppalgorithm}{RQR--DESN generalized-Bayes MCMC}
\emph{Input:} response vector \(\vect y\), fixed DESN design matrix
\(\mat X\), coverage level \(c\), learning rate \(\omega_{\mathrm R}\), coefficient prior,
initial roots \((\vect\beta_1,\vect\beta_2)\), burn-in, retained draws, and
thinning.

\begin{enumerate}[leftmargin=2.5em]
\item Initialize root coefficients and latent scales \(v_1,\ldots,v_T\).
\item For each MCMC iteration:
  \begin{enumerate}[leftmargin=2.5em]
  \item update each \(v_t\) from the \(\GIG\) full conditional using the
  current residual product \(e_t\);
  \item update \(\vect\beta_1\) from its Gaussian block conditional on
  \(\vect\beta_2\) and \(\vect v\);
  \item update the first-root shrinkage state when using \texttt{rhs\_ns};
  \item update \(\vect\beta_2\) from its Gaussian block conditional on
  \(\vect\beta_1\) and \(\vect v\);
  \item update the second-root shrinkage state when using \texttt{rhs\_ns};
  \item record the RQR loss and numerical precision diagnostics.
  \end{enumerate}
\item After burn-in and thinning, transform every retained draw to ordered
interval endpoints
\[
L_t^{(s)}=\min(\vect x_t^\top\vect\beta_1^{(s)},
\vect x_t^\top\vect\beta_2^{(s)}),\qquad
U_t^{(s)}=\max(\vect x_t^\top\vect\beta_1^{(s)},
\vect x_t^\top\vect\beta_2^{(s)}).
\]
\end{enumerate}

\emph{Output:} posterior draws and summaries of the ordered interval endpoints,
widths, and midpoints. Root labels are exchangeable, so reported summaries are
formed after ordering endpoints draw by draw.
\end{suppalgorithm}

The article-facing validation uses MCMC only. A fast VB implementation may be
useful for screening or initialization, but VB interval uncertainty is not
promoted here because it has not undergone a separate calibration study. The
reported simulation evidence therefore uses empirical coverage, interval score,
width, and paired baseline deltas for MCMC interval summaries.

\clearpage
\section{Supplementary Validation Tables}
\label{sec:supp_validation_tables}

The main article reports MCMC-only validation summaries. This section preserves
the generated VB--LD companion panels, balanced joint-validation provenance,
and full generated table captions so the computational comparison remains
auditable without making VB rows the article-facing validation results.

\subsection{Single-Quantile VB--LD Companion Panels}

The following generated tables retain both inference panels from the pinned
500-observation validation bundle. In the main article, only the MCMC rows from
these source tables are displayed.

\input{tables/qdesn_validation_tt500_final_tables.tex}

\clearpage
\subsection{Joint Multi-Quantile Balanced MCMC Provenance}

The following generated tables preserve the protocol, scenario-level balanced
MCMC summaries, gate audit, and metric-specific winners for the joint and
independent multi-quantile readouts. The VB and VB--LD stages remain the
screening, calibration, and initialization layer; the tables below document the
MCMC confirmation evidence used by the main article.

\input{tables/joint_qdesn_article_validation_provenance_tables.tex}

\clearpage
\section{GloFAS Discrepancy-Calibration Model}
\label{sec:glofas_discrepancy}

The streamflow application model uses one forecast system, the Global Flood
Awareness System (GloFAS), together with a reference gauge record
\citep{AlfieriEtAl2013GloFAS,HarriganEtAl2023GloFASForecasts}. The reference
quantile and the GloFAS discrepancy receive separate coefficient vectors. The
historical reservoir states are fixed after preprocessing and washout. At
issued forecast horizons, however, the reference path is unobserved at the
forecast origin, so future reservoir states may depend on latent future
reference values. This section records the fixed-design conditional
calculations used by the bridge model and by any latent-path sampler
conditional on a proposed future path. The full latent-path posterior also
contains the missing future reference values and the induced future reservoir
states.

\subsection{Data Blocks and Augmented Design}
\label{subsec:glofas_aug_design}

Fix a quantile level \(p_0\). Let \(y_t\) denote transformed reference
streamflow, \(g_t^{\mathrm{ret}}\) the retrospective GloFAS product, and
\(g_{T+h,m}^{\mathrm{ens}}\) GloFAS ensemble member \(m\) at forecast origin
\(T\) and horizon \(h\). Let \(\mathcal I_{\mathrm{ref}}\) index reference
observations and let \(\mathcal I_{\mathrm{glo}}\) index both retrospective
and issued-ensemble GloFAS values. For \(i\in\mathcal I_c\),
\(c\in\mathcal C=\{\mathrm{ref},\mathrm{glo}\}\), write the observed value as
\(z_i^c\). The reference feature vector
\(\vect x_i^{\mathrm{ref}}\) is generated from the reference-process history.
The discrepancy feature vector \(\vect x_i^{\mathrm{disc}}\) is generated from
the GloFAS-minus-reference discrepancy path. Both vectors are fixed after
conditioning on the relevant historical or latent future path.

Define
\[
\vect\theta_{p_0}
=
\begin{bmatrix}
\vect\beta_{p_0}\\
\vect\alpha_{p_0}
\end{bmatrix},
\qquad
\vect h_t^{\mathrm{ref}}=
\begin{bmatrix}
\vect x_t^{\mathrm{ref}}\\
\vect 0
\end{bmatrix},
\qquad
\vect h_t^{\mathrm{glo}}=
\begin{bmatrix}
\vect x_t^{\mathrm{ref}}\\
\vect x_t^{\mathrm{disc}}
\end{bmatrix}.
\]
The same definition applies to any indexed contribution \(i\). The
source-specific readout locations are
\[
q^{\mathrm{ref}}_{p_0,i}
=\vect h_i^{\mathrm{ref}\,\top}\vect\theta_{p_0}
=\vect x_i^{\mathrm{ref}\,\top}\vect\beta_{p_0},
\qquad
q^{\mathrm{glo}}_{p_0,i}
=\vect h_i^{\mathrm{glo}\,\top}\vect\theta_{p_0}
=\vect x_i^{\mathrm{ref}\,\top}\vect\beta_{p_0}
+\vect x_i^{\mathrm{disc}\,\top}\vect\alpha_{p_0}.
\]
Thus \(\vect\beta_{p_0}\) represents the calibrated reference quantile and
\(\vect\alpha_{p_0}\) represents the GloFAS discrepancy at the same level. In
the historical rows, \(\vect x_i^{\mathrm{disc}}\) is driven by
\(g_i^{\mathrm{ret}}-y_i\). In the current latent-path implementation for
issued forecast horizons, it is driven by the keyed path
\(\hat q^{\mathrm{ens}}_{p_0,T+h}-y_{T+h}\), where
\(\hat q^{\mathrm{ens}}_{p_0,T+h}\) is the empirical GloFAS ensemble quantile
at the fitted level.
Stacking source contributions conditional on a fixed state path gives a vector
\(\vect z\), a design matrix \(\mat H\), and source labels \(c_i\). With
observed historical data only, \(\mat H\) is fixed. In the latent-path
ensemble-likelihood model, the forecast-window rows of \(\mat H\) are functions
of the missing future reference path and must be recomputed inside inference.

The forecast-time use of this decomposition is draw-based and subtractive. For
horizons covered by an issued GloFAS ensemble at origin \(T\), the ensemble
informs the forecast-system quantile \(q^{\mathrm{glo}}_{p_0,T+h}\). For
posterior draw \(s\), the reference-process quantile draw is
\[
q^{\mathrm{ref},(s)}_{p_0,T+h}
=
q^{\mathrm{glo},(s)}_{p_0,T+h}-d^{\mathrm{glo},(s)}_{p_0,T+h}.
\]
Point summaries are taken after this subtraction and are not the primary
Bayesian prediction object. A posterior predictive draw for reference
streamflow can then be generated from the AL or exAL working likelihood using
the matched draws of the calibrated quantile, source-specific scale, and, for
exAL, asymmetry. Using future observed reference flows to construct
\(\vect x_{T+h}\) would leak validation information. The latent-path model
therefore treats those future reference values as missing and recomputes only
the \(H\) forecast-window state rows for each proposed future path. The
horizon-indexed origin-state row \((\vect x_T^\top,h/H)^\top\) remains a
documented bridge approximation and diagnostic baseline, not the final
ensemble-likelihood target.

\subsection{AL and exAL Working Likelihoods}
\label{subsec:glofas_likelihoods}

For exAL, each source \(c\in\mathcal C\) has scale and asymmetry parameters
\((\sigma_c,\gamma_c)\). Conditional on latent variables
\(\{v_i^c,s_i^c:i\in\mathcal I_c\}\),
\begin{align}
v_i^c\mid\sigma_c&\sim\Exp(\text{rate}=1/\sigma_c),&
s_i^c&\sim\TN(0,1),\\
z_i^c\mid\vect\theta_{p_0},\sigma_c,\gamma_c,v_i^c,s_i^c
&\sim
\Normal\left(
\vect h_i^{c\top}\vect\theta_{p_0}
+\sigma_cD_{\gamma_c}s_i^c+A_{\gamma_c}v_i^c,
\sigma_cB_{\gamma_c}v_i^c
\right).
\label{eq:supp_glofas_exal_hierarchy}
\end{align}
The marginal \(p_0\)-quantile of \(z_i^c\mid\vect h_i^c\) is
\(\vect h_i^{c\top}\vect\theta_{p_0}\). The AL working likelihood is obtained
by omitting \(s_i^c\) and \(\gamma_c\), replacing \(A_{\gamma_c}\) and
\(B_{\gamma_c}\) with \(A_0\) and \(B_0\), and using
\begin{equation}
z_i^c\mid\vect\theta_{p_0},\sigma_c,v_i^c
\sim
\Normal\left(
\vect h_i^{c\top}\vect\theta_{p_0}+A_0v_i^c,
\sigma_cB_0v_i^c
\right).
\label{eq:supp_glofas_al_hierarchy}
\end{equation}

The GloFAS block \(\mathcal I_{\mathrm{glo}}\) may contain both retrospective values and
issued ensemble members. Ensemble members at a common origin and horizon enter
as conditionally independent likelihood contributions given the forecast-system
quantile path and the GloFAS likelihood parameters. This is a likelihood
factorization motivated by the operational ensemble construction, in which
perturbed initial states and model perturbations generate independent forecast
alternatives \citep{ECMWF2026ENSGeneration}; it is not an unconditional
independence claim about all hydrologic forecast errors. If diagnostics show
that retrospective and issued-ensemble residual behavior differs materially,
\(\mathcal I_{\mathrm{glo}}\) can be split into two GloFAS source labels with separate
scale and asymmetry blocks.

\subsection{Default Regularized-Horseshoe Readout Prior}
\label{subsec:glofas_rhs_prior}

For the application derivation, the regularized horseshoe is the default
readout prior. Let \(\mathcal J_\beta\) and \(\mathcal J_\alpha\) denote the
non-intercept entries of \(\vect\beta_{p_0}\) and \(\vect\alpha_{p_0}\). The
reference and discrepancy intercepts are assigned weak Gaussian priors. For
block \(b\in\{\beta,\alpha\}\) and \(j\in\mathcal J_b\), write the corresponding
coefficient as \(\theta_{b,j}\). The product representation uses
\begin{equation}
\Normal(\theta_{b,j}\mid0,\tau_b^2\lambda_{b,j}^2)
\Normal(0\mid\theta_{b,j},\zeta_b^2).
\label{eq:supp_glofas_rhs_product}
\end{equation}
Conditional on the scales, this factor is proportional in \(\theta_{b,j}\) to a
Gaussian density with variance
\begin{equation}
V_{b,j}=\left(\zeta_b^{-2}+\tau_b^{-2}\lambda_{b,j}^{-2}\right)^{-1}.
\label{eq:supp_glofas_rhs_cond}
\end{equation}
The auxiliary representation for
\((\lambda_{b,j}^2,\nu_{b,j},\tau_b^2,\xi_b,\zeta_b^2)\) is the same as
\eqref{eq:supp_hs_aux}--\eqref{eq:supp_zeta_prior}, with
\(\beta_j\) replaced by \(\theta_{b,j}\) and \(r\) replaced by
\(|\mathcal J_b|\). The two RHS hierarchies are independent a priori, although
the Gaussian coefficient update couples \(\vect\beta_{p_0}\) and
\(\vect\alpha_{p_0}\) through the augmented likelihood. The source-specific
exAL scale and asymmetry blocks are the only non-conjugate blocks introduced by
the working likelihood.

\subsection{Complete-Data Posterior}
\label{subsec:glofas_posterior}

Let
\[
\Theta_b
=\{(\lambda_{b,j}^2,\nu_{b,j}):j\in\mathcal J_b\},\tau_b^2,\xi_b,\zeta_b^2,
\qquad b\in\{\beta,\alpha\},
\]
and let \(\vect v^c=(v_i^c:i\in\mathcal I_c)\) and
\(\vect s^c=(s_i^c:i\in\mathcal I_c)\). Under exAL,
\begin{align}
&p(\vect z,\{\vect v^c,\vect s^c,\sigma_c,\gamma_c\}_{c\in\mathcal C},
\vect\theta_{p_0},\Theta_\beta,\Theta_\alpha\mid\mat H)
\nonumber\\
&\quad\propto
\prod_{c\in\mathcal C}
\left[
\left\{
\prod_{i\in\mathcal I_c}
p(z_i^c\mid\vect\theta_{p_0},\sigma_c,\gamma_c,v_i^c,s_i^c,\vect h_i^c)
p(v_i^c\mid\sigma_c)p(s_i^c)
\right\}
p(\sigma_c)\pi_{\gamma_c}(\gamma_c)\ind\{L<\gamma_c<U\}
\right]
\nonumber\\
&\qquad\times
p(\vect\theta_{\mathrm{int}})
\prod_{b\in\{\beta,\alpha\}}
\Biggl[
\prod_{j\in\mathcal J_b}
\left\{
\Normal(\theta_{b,j}\mid0,\tau_b^2\lambda_{b,j}^2)
\Normal(0\mid\theta_{b,j},\zeta_b^2)
p(\lambda_{b,j}^2\mid\nu_{b,j})p(\nu_{b,j})
\right\}
\nonumber\\
&\qquad\qquad\times
p(\tau_b^2\mid\xi_b)p(\xi_b)p(\zeta_b^2)
\Biggr].
\label{eq:supp_glofas_joint_exal_rhs}
\end{align}
Here \(p(\vect\theta_{\mathrm{int}})\) denotes the weak Gaussian prior for the
two intercepts. Under AL,
\begin{align}
&p(\vect z,\{\vect v^c,\sigma_c\}_{c\in\mathcal C},
\vect\theta_{p_0},\Theta_\beta,\Theta_\alpha\mid\mat H)
\nonumber\\
&\quad\propto
\prod_{c\in\mathcal C}
\left[
\left\{
\prod_{i\in\mathcal I_c}
p(z_i^c\mid\vect\theta_{p_0},\sigma_c,v_i^c,\vect h_i^c)
p(v_i^c\mid\sigma_c)
\right\}
p(\sigma_c)
\right]
\nonumber\\
&\qquad\times
p(\vect\theta_{\mathrm{int}})
\prod_{b\in\{\beta,\alpha\}}
\Biggl[
\prod_{j\in\mathcal J_b}
\left\{
\Normal(\theta_{b,j}\mid0,\tau_b^2\lambda_{b,j}^2)
\Normal(0\mid\theta_{b,j},\zeta_b^2)
p(\lambda_{b,j}^2\mid\nu_{b,j})p(\nu_{b,j})
\right\}
\nonumber\\
&\qquad\qquad\times
p(\tau_b^2\mid\xi_b)p(\xi_b)p(\zeta_b^2)
\Biggr].
\label{eq:supp_glofas_joint_al_rhs}
\end{align}
The fixed-design posterior is the corresponding joint density viewed as a
function of unknown quantities conditional on \((\vect z,\mat H)\). In the
latent-path ensemble-likelihood model, \(\mat H\) should be read as
\(\mat H(\vect y_{T+1:T+H})\) for forecast-window rows, and the posterior is
augmented by the missing future reference path.
Forecast-window output inputs are strictly lagged: the design row for
\(T+h\) may depend on \(y_{T+h-1},y_{T+h-2},\ldots\), but not on
\(y_{T+h}\) itself.

\subsection{Full Conditional Posteriors}
\label{subsec:glofas_conditionals}

The full conditionals follow from the augmented Gaussian representation.
Define, for exAL,
\[
z_i^{c\star}
=z_i^c-\sigma_cD_{\gamma_c}s_i^c-A_{\gamma_c}v_i^c,\qquad
w_i^c=(\sigma_cB_{\gamma_c}v_i^c)^{-1}.
\]
For AL, use
\[
z_i^{c\star}=z_i^c-A_0v_i^c,\qquad
w_i^c=(\sigma_cB_0v_i^c)^{-1}.
\]
Stacking \(z_i^{c\star}\) and \(w_i^c\) over sources gives
\(\vect z^\star\) and \(\mat\Omega=\diag(w_i^c)\). Let
\(\mat P_\theta\) denote the conditional prior precision implied by the
weak Gaussian intercept priors and the RHS scales, with diagonal entries
\(\tau^{-2}\lambda_j^{-2}+\zeta^{-2}\) for \(j\in\mathcal J\). Let
\(\vect b_\theta\) collect the prior means, with zero entries for RHS-shrunk
coefficients. Then
\begin{align}
\mat\Sigma_{\theta\mid\cdot}
&=(\mat H^\top\mat\Omega\mat H+\mat P_\theta)^{-1},\\
\vect m_{\theta\mid\cdot}
&=\mat\Sigma_{\theta\mid\cdot}
(\mat H^\top\mat\Omega\vect z^\star+\mat P_\theta\vect b_\theta),\\
\vect\theta_{p_0}\mid\cdot
&\sim
\Normal(\vect m_{\theta\mid\cdot},\mat\Sigma_{\theta\mid\cdot}).
\label{eq:supp_glofas_theta_mcmc}
\end{align}

For exAL latent variables, define
\[
\delta_i^c
=z_i^c-\vect h_i^{c\top}\vect\theta_{p_0}
-\sigma_cD_{\gamma_c}s_i^c,
\qquad
\chi_i^c=\frac{(\delta_i^c)^2}{\sigma_cB_{\gamma_c}},
\qquad
\psi_i^c=\frac{A_{\gamma_c}^2}{\sigma_cB_{\gamma_c}}+\frac{2}{\sigma_c}.
\]
Then
\begin{equation}
v_i^c\mid\cdot\sim\GIG(1/2,\chi_i^c,\psi_i^c).
\label{eq:supp_glofas_v_mcmc}
\end{equation}
For \(s_i^c\), set
\(\mu_{0i}^c=z_i^c-\vect h_i^{c\top}\vect\theta_{p_0}
-A_{\gamma_c}v_i^c\). Then
\begin{align}
a_{s,i}^c&=1+\frac{\sigma_cD_{\gamma_c}^2}{B_{\gamma_c}v_i^c},&
m_{s,i}^c&=
\frac{D_{\gamma_c}\mu_{0i}^c}
{B_{\gamma_c}v_i^c+\sigma_cD_{\gamma_c}^2},\\
s_i^c\mid\cdot&\sim\TN(m_{s,i}^c,(a_{s,i}^c)^{-1}).
\label{eq:supp_glofas_s_mcmc}
\end{align}
For AL,
\begin{equation}
v_i^c\mid\cdot
\sim
\GIG\left(
1/2,
\frac{(z_i^c-\vect h_i^{c\top}\vect\theta_{p_0})^2}{\sigma_cB_0},
\frac{A_0^2}{\sigma_cB_0}+\frac{2}{\sigma_c}
\right).
\label{eq:supp_glofas_al_v_mcmc}
\end{equation}
If \(\sigma_c\sim\IG(a_{\sigma,c},b_{\sigma,c})\), then under AL
\begin{align}
\sigma_c\mid\cdot
&\sim
\IG\left(
a_{\sigma,c}+\frac{3n_c}{2},
b_{\sigma,c}+\sum_{i\in\mathcal I_c}v_i^c
+\frac12\sum_{i\in\mathcal I_c}
\frac{(z_i^c-\vect h_i^{c\top}\vect\theta_{p_0}-A_0v_i^c)^2}
{B_0v_i^c}
\right),
\label{eq:supp_glofas_al_sigma_mcmc}
\end{align}
where \(n_c=|\mathcal I_c|\).

For exAL, update \((\sigma_c,\gamma_c)\) source by source. The joint
log-kernel is
\begin{align}
\ell_c(\sigma_c,\gamma_c)
&=
-\frac12\sum_{i\in\mathcal I_c}\log(\sigma_cB_{\gamma_c}v_i^c)
-\frac12\sum_{i\in\mathcal I_c}
\frac{
\{z_i^c-\vect h_i^{c\top}\vect\theta_{p_0}
-\sigma_cD_{\gamma_c}s_i^c-A_{\gamma_c}v_i^c\}^2}
{\sigma_cB_{\gamma_c}v_i^c}
\nonumber\\
&\quad
-n_c\log\sigma_c-\sum_{i\in\mathcal I_c}\frac{v_i^c}{\sigma_c}
-(a_{\sigma,c}+1)\log\sigma_c-\frac{b_{\sigma,c}}{\sigma_c}
+\log\pi_{\gamma_c}(\gamma_c),
\label{eq:supp_glofas_sigmagam_kernel}
\end{align}
for \(L<\gamma_c<U\), with value \(-\infty\) outside this interval. This
kernel is the target for the source-specific scale and asymmetry update.

For each block \(b\in\{\beta,\alpha\}\) and \(j\in\mathcal J_b\), the RHS
scale updates are
\begin{align}
\lambda_{b,j}^2\mid\cdot
&\sim
\IG\left(1,\frac{1}{\nu_{b,j}}+\frac{\theta_{b,j}^2}{2\tau_b^2}\right),\\
\nu_{b,j}\mid\cdot
&\sim
\IG\left(1,1+\frac{1}{\lambda_{b,j}^2}\right),\\
\tau_b^2\mid\cdot
&\sim
\IG\left(\frac{|\mathcal J_b|+1}{2},
\frac{1}{\xi_b}+\frac12\sum_{j\in\mathcal J_b}\frac{\theta_{b,j}^2}{\lambda_{b,j}^2}\right),\\
\xi_b\mid\cdot
&\sim
\IG\left(1,\frac{1}{\tau_{0,b}^2}+\frac{1}{\tau_b^2}\right),\\
\zeta_b^2\mid\cdot
&\sim
\IG\left(a_{\zeta,b}+\frac{|\mathcal J_b|}{2},
b_{\zeta,b}+\frac12\sum_{j\in\mathcal J_b}\theta_{b,j}^2\right),
\label{eq:supp_glofas_rhs_mcmc}
\end{align}
with the final line omitted when the slab scale is fixed.

\subsection{MCMC Sampler}
\label{subsec:glofas_mcmc}

For exAL, use unconstrained coordinates
\[
\eta_{\sigma,c}=\log\sigma_c,\qquad
\eta_{\gamma,c}=\logit\{(\gamma_c-L)/(U-L)\}.
\]
The slice sampler targets
\begin{equation}
\ell_c\{T_c(\vect\eta_c)\}+\log|J_{T_c}(\vect\eta_c)|,
\qquad
\vect\eta_c=(\eta_{\gamma,c},\eta_{\sigma,c})^\top,
\label{eq:supp_glofas_slice_target}
\end{equation}
where \(T_c\) maps unconstrained coordinates to
\((\gamma_c,\sigma_c)\). This bivariate slice update may be replaced by a
Laplace-informed proposal or by one-dimensional conditional slice updates, but
the target density is \eqref{eq:supp_glofas_slice_target}.

\begin{suppalgorithm}{GloFAS discrepancy-calibration MCMC sampler}
\label{alg:glofas_mcmc}
\emph{Input:} stacked observations \(\vect z\), source labels \(c_i\), target
quantile \(p_0\), likelihood choice AL or exAL, RHS hyperparameters, MCMC
controls, and either a fixed augmented design \(\mat H\) or a latent-path
state-continuation map.
\begin{enumerate}[label=(\arabic*)]
\item Initialize \(\vect\theta_{p_0}\), RHS scales, source scales
\(\{\sigma_c\}\), latent \(\{v_i^c\}\), and, for exAL,
\(\{\gamma_c,s_i^c\}\). For the latent-path model, also initialize
\(\vect y_F\).
\item For each MCMC iteration:
  \begin{enumerate}[label=(\alph*)]
  \item for the latent-path model, rebuild the forecast-window rows of
  \(\mat H(\vect y_F)\);
  \item update each \(v_i^c\) from \eqref{eq:supp_glofas_v_mcmc} for exAL or
  \eqref{eq:supp_glofas_al_v_mcmc} for AL;
  \item for exAL, update each \(s_i^c\) from
  \eqref{eq:supp_glofas_s_mcmc};
  \item update \(\vect\theta_{p_0}\) from
  \eqref{eq:supp_glofas_theta_mcmc};
  \item update the beta and alpha RHS scales from \eqref{eq:supp_glofas_rhs_mcmc};
  \item for AL, update each \(\sigma_c\) from
  \eqref{eq:supp_glofas_al_sigma_mcmc};
  \item for exAL, update each \((\sigma_c,\gamma_c)\) by slice sampling on
  the transformed target \eqref{eq:supp_glofas_slice_target};
  \item for the latent-path model, update \(\vect y_F\) by a documented
  non-conjugate block or componentwise sampler, recomputing only the
  forecast-window state rows after a proposal;
  \item retain draws after burn-in, with thinning only if used.
  \end{enumerate}
\end{enumerate}
\end{suppalgorithm}

\subsection{Variational Bayes and CAVI Updates}
\label{subsec:glofas_vb}

For exAL, use the mean-field family
\begin{equation}
q=
q_\theta(\vect\theta_{p_0})
q_{\Theta_\beta}(\Theta_\beta)
q_{\Theta_\alpha}(\Theta_\alpha)
\prod_{c\in\mathcal C}
\left[
q_{\sigma_c,\gamma_c}(\sigma_c,\gamma_c)
\prod_{i\in\mathcal I_c}q_{v_i^c}(v_i^c)q_{s_i^c}(s_i^c)
\right]
\label{eq:supp_glofas_vb_family}
\end{equation}
where each \(q_{\Theta_b}\) factorizes over the auxiliary inverse-gamma RHS
components used in Section~\ref{subsec:glofas_rhs_prior}.
For AL, replace \(q_{\sigma_c,\gamma_c}\) by \(q_{\sigma_c}\) and omit
\(q_{s_i^c}\). The coefficient factor is Gaussian. Let
\[
\bar w_i^c=\E\{(\sigma_cB_{\gamma_c}v_i^c)^{-1}\},
\qquad
\bar m_i^c=
\E\left[
\frac{z_i^c-\sigma_cD_{\gamma_c}s_i^c-A_{\gamma_c}v_i^c}
{\sigma_cB_{\gamma_c}v_i^c}
\right]
\]
for exAL, with \(A_{\gamma_c},B_{\gamma_c},D_{\gamma_c}\) replaced by
\(A_0,B_0,0\) for AL. Stacking \(\bar w_i^c\) and \(\bar m_i^c\) gives
\(\bar{\mat\Omega}\) and \(\bar{\vect m}\). Let
\(\bar{\mat P}_\theta\) replace each block-specific RHS precision by
\(\E(\tau_b^{-2})\E(\lambda_{b,j}^{-2})+\E(\zeta_b^{-2})\). Then
\begin{align}
q_\theta(\vect\theta_{p_0})&=\Normal(\vect m_\theta,\mat\Sigma_\theta),\\
\mat\Sigma_\theta&=(\mat H^\top\bar{\mat\Omega}\mat H+\bar{\mat P}_\theta)^{-1},\\
\vect m_\theta&=\mat\Sigma_\theta(\mat H^\top\bar{\vect m}
+\bar{\mat P}_\theta\vect b_\theta).
\label{eq:supp_glofas_vb_theta}
\end{align}

For each contribution, define
\[
\bar r_i^c=z_i^c-\vect h_i^{c\top}\vect m_\theta,\qquad
\overline{(r_i^c)^2}
=(\bar r_i^c)^2+\vect h_i^{c\top}\mat\Sigma_\theta\vect h_i^c .
\]
For AL,
\begin{align}
q(v_i^c)
&=
\GIG\left(
1/2,
B_0^{-1}\E(\sigma_c^{-1})\overline{(r_i^c)^2},
\E(\sigma_c^{-1})\left(\frac{A_0^2}{B_0}+2\right)
\right),
\label{eq:supp_glofas_vb_al_v}\\
q(\sigma_c)
&=
\IG(a_{\sigma,c}^\star,b_{\sigma,c}^\star),\\
a_{\sigma,c}^\star
&=a_{\sigma,c}+\frac{3n_c}{2},\\
b_{\sigma,c}^\star
&=
b_{\sigma,c}+\sum_{i\in\mathcal I_c}\E(v_i^c)
+\frac{1}{2B_0}\sum_{i\in\mathcal I_c}
\left[
\overline{(r_i^c)^2}\E\{(v_i^c)^{-1}\}
-2A_0\bar r_i^c+A_0^2\E(v_i^c)
\right].
\label{eq:supp_glofas_vb_al_sigma}
\end{align}

For exAL, define source-specific smooth expectations
\[
M_{1c}=\E\{(\sigma_cB_{\gamma_c})^{-1}\},\quad
M_{2c}=\E(D_{\gamma_c}/B_{\gamma_c}),\quad
M_{3c}=\E(\sigma_cD_{\gamma_c}^2/B_{\gamma_c}),\quad
M_{4c}=\E(A_{\gamma_c}D_{\gamma_c}/B_{\gamma_c}),
\]
and
\[
M_{5c}=\E\{A_{\gamma_c}^2/(\sigma_cB_{\gamma_c})+2/\sigma_c\}.
\]
These are smooth expectations involving the source-specific exAL
scale and asymmetry factor, evaluated by Laplace--Delta. The latent factors are
\begin{align}
q(v_i^c)
&=
\GIG(1/2,\bar\chi_i^c,\bar\psi_i^c),\\
\bar\chi_i^c
&=
M_{1c}\overline{(r_i^c)^2}
-2M_{2c}\bar r_i^c\E(s_i^c)
+M_{3c}\E\{(s_i^c)^2\},\\
\bar\psi_i^c&=M_{5c},
\label{eq:supp_glofas_vb_exal_v}\\
q(s_i^c)&=\TN(m_{s,i}^{c,\mathrm{VB}},V_{s,i}^{c,\mathrm{VB}}),\\
V_{s,i}^{c,\mathrm{VB}}
&=\left[1+M_{3c}\E\{(v_i^c)^{-1}\}\right]^{-1},\\
m_{s,i}^{c,\mathrm{VB}}
&=
V_{s,i}^{c,\mathrm{VB}}
\left[
M_{2c}\bar r_i^c\E\{(v_i^c)^{-1}\}-M_{4c}
\right].
\label{eq:supp_glofas_vb_exal_s}
\end{align}

The RHS CAVI updates are applied independently to each block
\(b\in\{\beta,\alpha\}\):
\begin{align}
q(\lambda_{b,j}^2)
&=\IG\left(1,\ \E(\nu_{b,j}^{-1})
+\frac12\overline{\theta_{b,j}^2}\E(\tau_b^{-2})\right),\\
q(\nu_{b,j})
&=\IG\left(1,\ 1+\E(\lambda_{b,j}^{-2})\right),\\
q(\tau_b^2)
&=\IG\left(\frac{|\mathcal J_b|+1}{2},
\E(\xi_b^{-1})
+\frac12\sum_{j\in\mathcal J_b}\overline{\theta_{b,j}^2}
\E(\lambda_{b,j}^{-2})\right),\\
q(\xi_b)
&=\IG\left(1,\ \tau_{0,b}^{-2}+\E(\tau_b^{-2})\right),\\
q(\zeta_b^2)
&=\IG\left(a_{\zeta,b}+\frac{|\mathcal J_b|}{2},
b_{\zeta,b}+\frac12\sum_{j\in\mathcal J_b}\overline{\theta_{b,j}^2}\right),
\label{eq:supp_glofas_vb_rhs}
\end{align}
where
\(\overline{\theta_{b,j}^2}=m_{\theta,bj}^2+(\mat\Sigma_\theta)_{bj,bj}\).

For exAL, the source-specific Laplace--Delta target is
\begin{align}
\ell_{c,\mathrm{LD}}(\vect\eta_c)
&=
\E_{q_{-(\sigma_c,\gamma_c)}}\{
\log p(\vect z^c\mid\vect\theta_{p_0},\sigma_c,\gamma_c,
\vect v^c,\vect s^c,\{\vect h_i^c:i\in\mathcal I_c\})
\nonumber\\
&\qquad
+\log p(\vect v^c\mid\sigma_c)
+\log p(\sigma_c)+\log\pi_{\gamma_c}(\gamma_c)
\}
+\log|J_{T_c}(\vect\eta_c)|.
\label{eq:supp_glofas_ld_target}
\end{align}
At the mode \(\hat{\vect\eta}_c\), with Hessian
\(\mat K_c=\nabla^2\ell_{c,\mathrm{LD}}(\hat{\vect\eta}_c)\), set
\begin{equation}
q_{\sigma_c,\gamma_c}(\vect\eta_c)
\approx
\Normal\{\hat{\vect\eta}_c,(-\mat K_c)^{-1}\}.
\label{eq:supp_glofas_ld_gaussian}
\end{equation}
For any smooth function \(h(\sigma_c,\gamma_c)\),
\begin{equation}
\E\{h(\sigma_c,\gamma_c)\}
\approx
h\{T_c(\hat{\vect\eta}_c)\}
+\frac12\tr\left[
\nabla^2_\eta h\{T_c(\hat{\vect\eta}_c)\}(-\mat K_c)^{-1}
\right].
\label{eq:supp_glofas_delta}
\end{equation}

For the latent-path ensemble-likelihood model, the variational family is
augmented by a factor \(q_F(\vect y_F)\), where
\(\vect y_F=(y_{T+1},\ldots,y_{T+H})^\top\). Write
\(z_i^c(\vect y_F)=z_{i0}^c+\vect u_i^\top\vect y_F\), where
\(\vect u_i=\vect0\) for observed rows and selects the appropriate future
reference value for forecast-window reference rows. The forecast-window
feature row is \(\vect h_i^c(\vect y_F)\), because output lags in the
reservoir input may depend on \(\vect y_F\). For a current coefficient factor
\(q_\theta=\Normal(\vect m_\theta,\mat\Sigma_\theta)\), define
\[
e_i^c(\vect y_F)
=z_i^c(\vect y_F)-\vect h_i^c(\vect y_F)^\top\vect m_\theta,
\qquad
R_i^c(\vect y_F)
=\{e_i^c(\vect y_F)\}^2+
\vect h_i^c(\vect y_F)^\top
\mat\Sigma_\theta\vect h_i^c(\vect y_F).
\]
For AL, the future-path Laplace target is, up to constants not depending on
\(\vect y_F\),
\begin{equation}
\ell_F^{\mathrm{AL}}(\vect y_F)
=
-\frac{1}{2B_0}\sum_{i\in\mathcal I_F}
\E(\sigma_{c_i}^{-1})
\left[
\E\{(v_i^{c_i})^{-1}\}R_i^{c_i}(\vect y_F)
-2A_0e_i^{c_i}(\vect y_F)
\right],
\label{eq:supp_glofas_latent_path_al_ld_target}
\end{equation}
where \(\mathcal I_F\) contains rows whose response or feature row depends on
the future path. Terms that do not depend on \(\vect y_F\), including the
\(A_0^2\E(v_i^{c_i})\) part of the Gaussian likelihood and the exponential
latent-prior terms, are handled in the complete-data ELBO but do not enter the
future-path mode or Hessian. For exAL, the corresponding target is
\begin{align}
\ell_F^{\mathrm{exAL}}(\vect y_F)
&=
-\frac12\sum_{i\in\mathcal I_F}
\left[
M_{1,c_i}\E\{(v_i^{c_i})^{-1}\}R_i^{c_i}(\vect y_F)
-2M_{2,c_i}\E(s_i^{c_i})\E\{(v_i^{c_i})^{-1}\}
e_i^{c_i}(\vect y_F)
-2M_{6,c_i}e_i^{c_i}(\vect y_F)
\right],
\label{eq:supp_glofas_latent_path_exal_ld_target}
\end{align}
with
\[
M_{6,c}=\E\{A_{\gamma_c}/(\sigma_cB_{\gamma_c})\}.
\]
The remaining exAL terms in the complete-data likelihood and latent-prior
density do not depend on \(\vect y_F\) and are included in the ELBO
decomposition below.
At the mode \(\hat{\vect y}_F\), set
\[
\mat K_F=-\nabla^2_{\vect y_F}\ell_F(\hat{\vect y}_F),
\qquad
q_F(\vect y_F)\approx
\Normal(\hat{\vect y}_F,\mat K_F^{-1}),
\]
with any Hessian regularization recorded in the fit object. Delta-method
moments of the nonlinear state map are then used in the remaining coordinate
updates. If \(\mat J_i\) is the Jacobian of \(\vect h_i^c(\vect y_F)\) at
\(\hat{\vect y}_F\), and \(\mat V_F=\mat K_F^{-1}\), then the default
first-order Delta moments are
\[
\vect a_i^c
\approx
\vect h_i^c(\hat{\vect y}_F),
\qquad
\mat S_i^c
\approx
\vect a_i^c\vect a_i^{c\top}
+\mat J_i\mat V_F\mat J_i^\top .
\]
Also,
\[
\bar z_i^c=z_{i0}^c+\vect u_i^\top\hat{\vect y}_F,
\qquad
\overline{(z_i^c)^2}
=(\bar z_i^c)^2+\vect u_i^\top\mat V_F\vect u_i,
\qquad
\vect b_i^c
\approx
\vect a_i^c\bar z_i^c+\mat J_i\mat V_F\vect u_i .
\]
Second-order mean corrections can be added for selected low-dimensional
summaries, but they are not required for the first implementation because they
would require Hessians for all reservoir features.
Replacing fixed-design products by
\((\vect a_i^c,\mat S_i^c,\vect b_i^c,\bar z_i^c,
\overline{(z_i^c)^2})\) gives the latent-path VB--LD updates. The entropy
contribution of the future-path factor is
\[
H(q_F)=\frac12\log |2\pi e\,\mat V_F|.
\]
The resulting monitored objective is a Laplace--Delta approximation to the
ELBO of the nonlinear latent-path model.

\begin{suppalgorithm}{GloFAS discrepancy-calibration variational Bayes updates}
\label{alg:glofas_vb}
\emph{Input:} stacked observations \(\vect z\), source labels \(c_i\), target
quantile \(p_0\), likelihood choice AL or exAL, block-specific RHS hyperparameters,
convergence controls, and either a fixed augmented design \(\mat H\) or a
latent-path state-continuation map.
\begin{enumerate}[label=(\arabic*)]
\item Initialize \(q_\theta\), latent factors, source scale factors, and RHS
scale factors. For the latent-path model, also initialize \(q_F\).
\item Repeat until convergence:
  \begin{enumerate}[label=(\alph*)]
  \item for the latent-path model, update \(q_F\) by the Laplace target
  \eqref{eq:supp_glofas_latent_path_al_ld_target} for AL or
  \eqref{eq:supp_glofas_latent_path_exal_ld_target} for exAL, then recompute
  the Delta-method design moments;
  \item update \(q(v_i^c)\) by \eqref{eq:supp_glofas_vb_al_v} for AL or
  \eqref{eq:supp_glofas_vb_exal_v} for exAL;
  \item for exAL, update \(q(s_i^c)\) by
  \eqref{eq:supp_glofas_vb_exal_s};
  \item update \(q_\theta\) by \eqref{eq:supp_glofas_vb_theta};
  \item update the beta and alpha RHS factors by \eqref{eq:supp_glofas_vb_rhs};
  \item for AL, update each \(q(\sigma_c)\) by
  \eqref{eq:supp_glofas_vb_al_sigma};
  \item for exAL, update each \(q_{\sigma_c,\gamma_c}\) by
  \eqref{eq:supp_glofas_ld_target}--\eqref{eq:supp_glofas_ld_gaussian};
  \item recompute smooth expectations by \eqref{eq:supp_glofas_delta} and
  monitor the ELBO, or the Laplace--Delta ELBO approximation when \(q_F\) is
  present.
  \end{enumerate}
\item Return variational factors and summaries of
\(\vect h_i^{c\top}\vect\theta_{p_0}\).
\end{enumerate}
\end{suppalgorithm}

\subsection{ELBO}
\label{subsec:glofas_elbo}

For exAL, the application ELBO is
\begin{align}
\elbo(q)
&=
\sum_{c\in\mathcal C}
\left[
\E_q\{\log p(\vect z^c\mid\vect\theta_{p_0},\sigma_c,\gamma_c,
\vect v^c,\vect s^c,\mat H)\}
+\E_q\{\log p(\vect v^c\mid\sigma_c)\}
+\E_q\{\log p(\vect s^c)\}
\right.
\nonumber\\
&\qquad\left.
+\E_q\{\log p(\sigma_c)\}
+\E_q\{\log\pi_{\gamma_c}(\gamma_c)\}
\right]
+\E_q\{\log p(\vect\theta_{p_0},\Theta_\beta,\Theta_\alpha)\}
+H(q).
\label{eq:supp_glofas_elbo_decomp}
\end{align}
For AL, omit the \(s\)-latent and \(\gamma_c\) prior terms. The exAL
likelihood term for source \(c\) is
\begin{align}
\elbo_{\mathrm{like},c}
&=
-\frac{n_c}{2}\log(2\pi)
-\frac12\sum_{i\in\mathcal I_c}
\left[
\E(\log\sigma_c)+\E(\log B_{\gamma_c})+\E(\log v_i^c)
\right]
\nonumber\\
&\quad
-\frac12\sum_{i\in\mathcal I_c}
\E\left[
\frac{
\{z_i^c-\vect h_i^{c\top}\vect\theta_{p_0}
-\sigma_cD_{\gamma_c}s_i^c-A_{\gamma_c}v_i^c\}^2}
{\sigma_cB_{\gamma_c}v_i^c}
\right].
\label{eq:supp_glofas_elbo_like}
\end{align}
The corresponding AL likelihood term is
\begin{align}
\elbo_{\mathrm{like},c}^{\mathrm{AL}}
&=
-\frac{n_c}{2}\log(2\pi)
-\frac12\sum_{i\in\mathcal I_c}
\left[
\E(\log\sigma_c)+\log B_0+\E(\log v_i^c)
\right]
\nonumber\\
&\quad
-\frac12\sum_{i\in\mathcal I_c}
\E\left[
\frac{
\{z_i^c-\vect h_i^{c\top}\vect\theta_{p_0}-A_0v_i^c\}^2}
{\sigma_cB_0v_i^c}
\right].
\label{eq:supp_glofas_elbo_like_al}
\end{align}
For the latent-path model, these likelihood terms are evaluated with the
Laplace--Delta future-path moments above, and \(H(q)\) includes
\(H(q_F)\). Thus the monitored quantity is an ELBO approximation for the
nonlinear latent-path model; it reduces to the fixed-design ELBO when
\(\vect y_F\) is absent or held fixed.
The latent and source-parameter prior terms are
\begin{align}
\elbo_{v,c}
&=
-n_c\E(\log\sigma_c)-\E(\sigma_c^{-1})
\sum_{i\in\mathcal I_c}\E(v_i^c),\\
\elbo_{s,c}
&=
n_c\log2-\frac{n_c}{2}\log(2\pi)
-\frac12\sum_{i\in\mathcal I_c}\E\{(s_i^c)^2\},\\
\elbo_{\sigma,c}
&=
a_{\sigma,c}\log b_{\sigma,c}-\log\Gamma(a_{\sigma,c})
-(a_{\sigma,c}+1)\E(\log\sigma_c)-b_{\sigma,c}\E(\sigma_c^{-1}),\\
\elbo_{\gamma,c}
&=\E\{\log\pi_{\gamma_c}(\gamma_c)\}.
\label{eq:supp_glofas_elbo_source_terms}
\end{align}
The RHS coefficient-prior and shrinkage-scale terms are blockwise copies of
Section~\ref{subsec:rhs_elbo}, with \(\vect\beta\) replaced separately by the
non-intercept entries of \(\vect\beta_{p_0}\) and
\(\vect\alpha_{p_0}\). Entropy terms are summed over the Gaussian coefficient
factor, GIG latent factors, truncated-normal latent factors, inverse-gamma RHS
factors, and the source-specific Laplace--Delta factors.
Terms involving smooth nonlinear functions of \((\sigma_c,\gamma_c)\) are
evaluated with the same Laplace--Delta approximation used in
\eqref{eq:supp_glofas_delta}.

\subsection{Internal Checks}
\label{subsec:glofas_checks}

The derivation has several useful reductions. If the GloFAS source is removed
and \(\vect\alpha_{p_0}\) is omitted, the model reduces to the ordinary
Q--DESN readout in Sections~\ref{sec:model_joint}--\ref{sec:elbo}. If
\(\gamma_c=0\) and the \(s_i^c\) variables are omitted, the exAL updates reduce
to the AL updates with constants \(A_0\) and \(B_0\). The coefficient update
remains Gaussian because the augmented likelihood and the RHS prior are
conditionally Gaussian. The RHS hierarchy contributes only inverse-gamma scale
updates under the product representation; it does not create the
source-specific non-conjugate scale and asymmetry block. Finally, the MCMC
sampler and the variational approximation target the same complete-data joint
density, so the ELBO in \eqref{eq:supp_glofas_elbo_decomp} should use the same
factors as the MCMC posterior in \eqref{eq:supp_glofas_joint_exal_rhs}.

\bibliographystyle{apalike}
\bibliography{refs}

\end{document}
